% ============================================================
%  MIVA-KNIGHT-AGENT: Agentic Multimodal Retrieval-Augmented
%  Generation Voice Assistant with Autonomous Planning,
%  Tool-Use Orchestration, and Self-Reflective Reasoning
%
%  Full IEEE-style Research Paper — LaTeX Source
%  Builds upon MIVA-KNIGHT+ (LLM/VLM/LVM/MLM integration)
%  and adds a complete Agentic AI layer.
% ============================================================
\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts

%% ── Core packages ─────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{amsthm}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric,arrows.meta,positioning,fit,
                calc,backgrounds,decorations.pathreplacing,
                matrix,chains,scopes,shadows,
                decorations.markings,patterns}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{url}
\usepackage{cite}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabularx}
\usepackage{balance}
\usepackage{microtype}
\usepackage{bm}
\usepackage{bbm}
\usepackage{enumitem}
\usepackage{pifont}

%% ── Theorem environments (IEEEtran-compatible) ────────────
\newtheoremstyle{ieeethm}{4pt}{4pt}{\itshape}{0pt}{\bfseries}{.}{0.5em}{}
\newtheoremstyle{ieeedef}{4pt}{4pt}{\normalfont}{0pt}{\bfseries}{.}{0.5em}{}
\newtheoremstyle{ieeeremark}{4pt}{4pt}{\normalfont}{0pt}{\itshape}{.}{0.5em}{}
\theoremstyle{ieeethm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{ieeedef}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{axiom}{Axiom}
\theoremstyle{ieeeremark}
\newtheorem*{remark}{Remark}

%% ── Extended colour palette ───────────────────────────────
\definecolor{colLLM}{RGB}{34,85,170}
\definecolor{colVLM}{RGB}{160,40,120}
\definecolor{colLVM}{RGB}{20,140,80}
\definecolor{colMLM}{RGB}{200,100,20}
\definecolor{colAgent}{RGB}{80,30,140}      % deep purple — agent core
\definecolor{colPlan}{RGB}{0,120,180}       % planning module
\definecolor{colTool}{RGB}{180,80,0}        % tool executor
\definecolor{colMem}{RGB}{30,150,100}       % memory store
\definecolor{colRefl}{RGB}{200,40,60}       % reflector/critic
\definecolor{colSafe}{RGB}{100,100,100}     % safety guard
\definecolor{colFusion}{RGB}{120,60,160}
\definecolor{colOrch}{RGB}{60,100,30}       % orchestrator

%% ── TikZ global node styles ───────────────────────────────
\tikzset{
  %% generic
  block/.style   = {rectangle, rounded corners=3pt, draw=black!70,
                    fill=blue!8, text centered, minimum height=0.52cm,
                    minimum width=1.75cm, font=\scriptsize},
  sblock/.style  = {rectangle, rounded corners=3pt, draw=black!60,
                    fill=green!10, text centered, minimum height=0.48cm,
                    minimum width=1.55cm, font=\scriptsize},
  mblock/.style  = {rectangle, rounded corners=3pt, draw=colFusion!80,
                    fill=colFusion!8, text centered, minimum height=0.52cm,
                    minimum width=1.75cm, font=\scriptsize},
  rblock/.style  = {rectangle, rounded corners=3pt, draw=red!60,
                    fill=red!6, text centered, minimum height=0.52cm,
                    minimum width=1.55cm, font=\scriptsize},
  oblock/.style  = {rectangle, rounded corners=3pt, draw=orange!70,
                    fill=orange!8, text centered, minimum height=0.52cm,
                    minimum width=1.75cm, font=\scriptsize},
  dblock/.style  = {diamond, draw=black!70, fill=yellow!12,
                    text centered, minimum height=0.58cm,
                    minimum width=1.35cm, font=\tiny, aspect=2.5},
  %% foundation-model family blocks
  llmblock/.style= {rectangle, rounded corners=3pt, draw=colLLM,
                    fill=colLLM!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  vlmblock/.style= {rectangle, rounded corners=3pt, draw=colVLM,
                    fill=colVLM!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  lvmblock/.style= {rectangle, rounded corners=3pt, draw=colLVM,
                    fill=colLVM!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  mlmblock/.style= {rectangle, rounded corners=3pt, draw=colMLM,
                    fill=colMLM!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  %% agentic-layer blocks
  agentblock/.style={rectangle, rounded corners=5pt, draw=colAgent,
                    fill=colAgent!10, text centered, minimum height=0.58cm,
                    minimum width=2.1cm, font=\scriptsize\bfseries, very thick},
  planblock/.style= {rectangle, rounded corners=4pt, draw=colPlan,
                    fill=colPlan!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  toolblock/.style= {rectangle, rounded corners=4pt, draw=colTool,
                    fill=colTool!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  memblock/.style = {rectangle, rounded corners=4pt, draw=colMem,
                    fill=colMem!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  reflblock/.style= {rectangle, rounded corners=4pt, draw=colRefl,
                    fill=colRefl!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  safeblock/.style= {rectangle, rounded corners=4pt, draw=colSafe,
                    fill=colSafe!8, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  orchblock/.style= {rectangle, rounded corners=4pt, draw=colOrch,
                    fill=colOrch!10, text centered, minimum height=0.52cm,
                    minimum width=1.95cm, font=\scriptsize, thick},
  %% arrows
  arr/.style   = {-{Stealth[length=4pt]}, thick, draw=black!70},
  darr/.style  = {-{Stealth[length=4pt]}, thick, draw=colLLM!80},
  varr/.style  = {-{Stealth[length=4pt]}, thick, draw=colVLM!80},
  larr/.style  = {-{Stealth[length=4pt]}, thick, draw=colLVM!80},
  marr/.style  = {-{Stealth[length=4pt]}, thick, draw=colMLM!80},
  aarr/.style  = {-{Stealth[length=5pt]}, very thick, draw=colAgent!80},
  parr/.style  = {-{Stealth[length=4pt]}, thick, draw=colPlan!80},
  tarr/.style  = {-{Stealth[length=4pt]}, thick, draw=colTool!80},
  rarr/.style  = {-{Stealth[length=4pt]}, thick, draw=colRefl!80},
  %% feedback loop arrow style
  looparr/.style = {-{Stealth[length=4pt]}, thick, draw=colAgent!70,
                    dashed},
}

%% ── Column helper ─────────────────────────────────────────
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\algrenewcommand\algorithmicindent{0.8em}

%% ── Hyperref (load last) ──────────────────────────────────
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!60!black,
            citecolor=blue!60!black,urlcolor=blue!60!black}

% ============================================================
\begin{document}

\title{MIVA-KNIGHT-AGENT: An Agentic Multimodal\\
Retrieval-Augmented Generation Voice Assistant\\
with Autonomous Planning, Tool Orchestration,\\
and Self-Reflective Reasoning}

\author{%
\IEEEauthorblockN{Oluwakayode Soyinka}
\IEEEauthorblockA{Department of Information Technology\\
Concordia University of Edmonton\\
Edmonton, AB, Canada\\
\texttt{kayode.soyinka@concordia.ab.ca}}
\and
\IEEEauthorblockN{Baidya Nath Saha}
\IEEEauthorblockA{Department of Information Technology\\
Concordia University of Edmonton\\
Edmonton, AB, Canada\\
\texttt{bsaha@concordia.ab.ca}}
}

\maketitle

%% ── Abstract ──────────────────────────────────────────────
\begin{abstract}
We present \textsc{MIVA-KNIGHT-AGENT} (\textbf{M}ultimodal \textbf{I}ntelligent \textbf{V}oice \textbf{A}ssistant --- \textbf{K}nowledge-\textbf{N}exus \textbf{I}ntegrated \textbf{G}eneration with \textbf{H}ybrid \textbf{T}ransformer and \textbf{A}gentic \textbf{G}eneration \textbf{E}ngine \textbf{N}exus \textbf{T}echnology), a next-generation agentic multimodal retrieval-augmented generation (MM-AgRAG) voice assistant that augments the \textsc{MIVA-KNIGHT}$^+$ foundation-model stack (LLM, VLM, LVM, MLM) with a full \emph{Agentic Intelligence Layer} (AIL). The AIL comprises five coordinated modules: (1) a \emph{Hierarchical Task Planner} (HTP) that decomposes complex multi-step queries into sub-goal trees using Monte Carlo Tree Search (MCTS)-guided LLM planning; (2) a \emph{Tool-Use Orchestrator} (TUO) that dynamically selects and sequences fifteen heterogeneous tools (retrieval, computation, summarisation, image captioning, NER, medical parsing, sentiment scoring, emotion detection, web search, SQL, code execution, translation, speech, TTS, and calendar) using a learned tool-selection policy $\pi_\theta$; (3) a \emph{Structured Memory Subsystem} (SMS) with episodic, semantic, and procedural stores backed by a persistent FAISS index and a Redis key--value cache; (4) a \emph{Self-Reflective Critic} (SRC) that scores each generated sub-answer with a fine-grained rubric (Correctness, Groundedness, Completeness, Safety) and triggers re-planning when any dimension falls below threshold; and (5) a \emph{Constitutional Safety Guard} (CSG) that enforces domain-specific safety constraints via a learned constraint classifier, preventing harmful generation in medical and general domains alike. The five \textsc{MIVA-KNIGHT}$^+$ modality pipelines (COCO, ROCO, RAVDESS, CREMA-D, CMU-MOSI) are preserved and enriched: the TUO can invoke any pipeline tool as an atomic action; the HTP routes domain-adaptive sub-goals to the ROCO or COCO FAISS index; the SRC uses SupCon emotion embeddings as affective-grounding signals. On agentic benchmarks, \textsc{MIVA-KNIGHT-AGENT} achieves: task completion rate of $84.6\%$ on a 200-query multi-step evaluation set; ROCO P@1 $= 71.3\%$ (vs. $68.9\%$ for non-agentic); CMU-MOSI Acc-2 $= 83.2\%$; CREMA-D emotion accuracy $= 54.7\%$; and constitutional safety constraint satisfaction $= 99.1\%$. The entire system operates at zero recurring API cost, demonstrating that research-grade agentic multimodal AI is achievable with open-weight models and commodity hardware.
\end{abstract}

\begin{IEEEkeywords}
Agentic AI, autonomous planning, tool-use orchestration, self-reflective reasoning, multimodal RAG, LLM agents, hierarchical task planning, Monte Carlo Tree Search, constitutional AI, memory-augmented agents, multi-agent systems, emotion-aware generation, knowledge graph, contrastive learning, voice assistant
\end{IEEEkeywords}

%%===================================================================
\section{Introduction}
\label{sec:intro}
%%===================================================================

The emergence of \emph{agentic artificial intelligence}---systems capable of autonomous goal decomposition, persistent memory, tool invocation, and self-critique---marks a qualitative shift beyond the reactive question-answering paradigm of classical retrieval-augmented generation (RAG)~\cite{lewis2020retrieval}. Where a non-agentic system maps a single query $x$ to a single response $r$, an agentic system pursues a \emph{goal} $\mathcal{G}$ across an unbounded sequence of actions, re-evaluating its own outputs and revising its plan in light of evidence gathered at each step.

Real-world voice assistants epitomise the need for agenticism. A user asking ``\emph{What is the most likely diagnosis for a 45-year-old patient with the chest X-ray I just uploaded, and what are the three most recent clinical trials addressing it?}'' is not issuing a retrieval query---they are commissioning a multi-step reasoning task that requires: (i) image analysis via a Large Vision Model (LVM), (ii) entity extraction via a Masked Language Model (MLM), (iii) medical document retrieval via a domain-adapted FAISS index, (iv) web search for trial registrations, (v) synthesis via a Large Language Model (LLM), and (vi) safety checking before spoken delivery via text-to-speech. No single-pass RAG pipeline can execute this reliably; an agent can.

\textsc{MIVA-KNIGHT}$^+$~\cite{soyinka2024mivaplus} provided the foundation: five modality pipelines (COCO, ROCO, RAVDESS, CREMA-D, CMU-MOSI), a shared 512-d contrastive embedding space, three-branch hybrid retrieval (FAISS + VLM VGA + NER-PMI KG), and four foundation-model families (LLM, VLM, LVM, MLM). \textsc{MIVA-KNIGHT-AGENT} wraps this stack inside a principled \emph{Agentic Intelligence Layer} (AIL), formalising the agent as a Markov Decision Process (MDP) whose actions are drawn from a structured tool library and whose policy is guided by an LLM planner with MCTS lookahead.

\textbf{Core agentic challenges addressed.}
\begin{itemize}[leftmargin=*,nosep]
  \item \textbf{Long-horizon planning:} decomposing queries into sub-goal trees with backtracking.
  \item \textbf{Tool-use generalisation:} selecting the right tool from a library of 15 heterogeneous actions.
  \item \textbf{Hallucination control:} a self-reflective critic that re-triggers retrieval when groundedness drops.
  \item \textbf{Affective grounding:} injecting emotion and sentiment signals as soft constraints on generation style.
  \item \textbf{Constitutional safety:} enforcing medical and general domain constraints without fine-tuning.
  \item \textbf{Memory persistence:} episodic recall of prior conversations and semantic caching of retrieved facts.
\end{itemize}

\textbf{Contributions.} This paper makes eight primary contributions:

\begin{enumerate}[leftmargin=*,nosep]
  \item We formalise \textsc{MIVA-KNIGHT-AGENT} as a Goal-Conditioned Partially Observable MDP (GC-POMDP) with a hybrid discrete-continuous action space covering modality encoding, retrieval, generation, and safety operations.

  \item We introduce the \emph{Hierarchical Task Planner} (HTP), which uses LLM-guided MCTS to search a sub-goal tree of depth $\leq 5$, provably converging to an $\epsilon$-optimal plan under bounded computational budget.

  \item We introduce the \emph{Tool-Use Orchestrator} (TUO) with a learned tool-selection policy $\pi_\theta$ trained via offline imitation learning on 5,000 expert-annotated tool-use traces, achieving $91.2\%$ tool selection accuracy on held-out queries.

  \item We introduce the \emph{Structured Memory Subsystem} (SMS) with episodic, semantic, and procedural tiers, formally proving that semantic cache hits reduce expected retrieval latency by a factor of $O(\sqrt{N})$ over full FAISS search.

  \item We introduce the \emph{Self-Reflective Critic} (SRC) with a four-dimension rubric (Correctness, Groundedness, Completeness, Safety) and prove that iterative self-critique with a bounded critic converges to a fixed-point answer within $K_\text{max} \leq \lceil \log_\mu(1/\epsilon) \rceil$ rounds.

  \item We introduce the \emph{Constitutional Safety Guard} (CSG), a lightweight (3M-parameter) constraint classifier trained on 12,000 constraint-annotated medical and general domain examples, achieving $99.1\%$ constraint satisfaction rate.

  \item We propose an \emph{Agentic Emotion Loop} (AEL) that feeds Pipeline~C/D emotion embeddings as affective state tokens into the HTP planning context, dynamically adjusting response style (empathetic, clinical, neutral) based on detected user affect.

  \item We demonstrate that the complete agentic stack---built entirely from open-weight models, local compute, and zero-cost APIs---achieves $84.6\%$ task completion on a new 200-query multi-step agentic benchmark.
\end{enumerate}

%%===================================================================
\section{Related Work}
\label{sec:related}
%%===================================================================

\subsection{Agentic AI and LLM-Based Agents}

The formalisation of LLMs as autonomous agents was catalysed by ReAct~\cite{yao2023react}, which interleaves \emph{Reason} and \emph{Act} steps in a scratchpad. Toolformer~\cite{schick2023toolformer} showed that LLMs can self-supervise tool-call insertion. AutoGPT~\cite{significant2023autogpt} and BabyAGI demonstrated long-horizon task loops but suffered from error accumulation without structured self-critique. Voyager~\cite{wang2023voyager} introduced a skill library for Minecraft agents using code generation. OpenAgents~\cite{xie2023openagents} provided a three-agent platform (data, plugin, web) for real-world tasks. Our TUO unifies these ideas into a learnable policy over 15 domain-specific tools, grounded in multimodal retrieval.

\subsection{Multi-Agent Orchestration}

MetaGPT~\cite{hong2023metagpt} assigns specialised roles (product manager, engineer, tester) to separate LLM agents coordinated by a shared message bus. AutoGen~\cite{wu2023autogen} enables arbitrary LLM-to-LLM conversation graphs. Our system differs in that a \emph{single} orchestrator agent drives all sub-modules via the HTP, with multi-agent parallelism used only for the SMS semantic cache and the CSG safety check to minimise latency.

\subsection{Self-Reflection and Critic Modules}

Reflexion~\cite{shinn2023reflexion} stores verbal self-reflection in episodic memory and shows improvement on coding and decision-making benchmarks. Self-Refine~\cite{madaan2023selfrefine} iterates feedback and refinement for open-ended generation. Our SRC is more structured: it scores along four explicit rubric dimensions and halts if the safety dimension fails, preventing the critic loop from eroding safety constraints.

\subsection{Planning in Agentic Systems}

Tree of Thoughts (ToT)~\cite{yao2023tot} systematically explores a tree of reasoning paths via BFS/DFS. Graph of Thoughts (GoT)~\cite{besta2023got} generalises to arbitrary DAGs. Reasoning via Planning (RAP)~\cite{hao2023rap} uses MCTS with an LLM as both world model and policy. Our HTP inherits the MCTS backbone from RAP and extends it with a \emph{multi-modal value estimator} that incorporates VLM image-query similarity and emotion state.

\subsection{Memory-Augmented Agents}

MemGPT~\cite{packer2023memgpt} manages finite context windows via tiered memory paging. HippoRAG~\cite{gutierrez2024hipporag} builds hippocampus-inspired knowledge graphs from retrieved passages. Our SMS adopts a three-tier architecture (episodic $\to$ semantic $\to$ procedural) consistent with cognitive memory models~\cite{atkinson1968human} and provides formal latency guarantees absent from prior work.

\subsection{Constitutional and Safe AI}

Constitutional AI (CAI)~\cite{bai2022constitutional} trains a critique-revision loop via a written constitution. RLHF-based safety~\cite{ouyang2022training} fine-tunes LLMs with human preference signals. Our CSG is a lightweight classifier (no fine-tuning required) that intercepts any generation whose constraint score $c < \theta_\text{safe}$, providing deterministic safety enforcement independent of the LLM's behaviour.

\subsection{Multimodal RAG and Foundation Models}

\textsc{MIVA-KNIGHT}$^+$~\cite{soyinka2024mivaplus} provides the multimodal backbone. FLAVA~\cite{singh2022flava}, UnifiedIO~\cite{lu2022unified}, and LLaVA~\cite{liu2023llava} address multimodal alignment. AudioCLIP~\cite{guzhov2022audioclip} and CLAP~\cite{elizalde2023clap} extend contrastive alignment to audio. Our work is the first to integrate a full agentic loop over all four foundation-model families in a zero-cost, single-system architecture.

%%===================================================================
\section{Formal Agent Model}
\label{sec:agent_model}
%%===================================================================

\subsection{Goal-Conditioned POMDP Formulation}

\begin{definition}[GC-POMDP Agent]
\label{def:gcpomdp}
The \textsc{MIVA-KNIGHT-AGENT} is formalised as a tuple $\mathcal{A} = \langle \mathcal{S}, \mathcal{O}, \mathcal{A}ct, \mathcal{G}, T, R, \Omega, \pi \rangle$ where:
\begin{itemize}[nosep]
  \item $\mathcal{S}$: latent world state (conversation history, retrieved facts, emotional state, domain).
  \item $\mathcal{O}$: observations (transcribed query, image pixels, retrieved documents, tool outputs).
  \item $\mathcal{A}ct = \mathcal{A}ct_\text{ret} \cup \mathcal{A}ct_\text{gen} \cup \mathcal{A}ct_\text{mem} \cup \mathcal{A}ct_\text{safe}$: hybrid action space of 15 atomic tool actions.
  \item $\mathcal{G}$: goal description, a natural-language string or multimodal tuple.
  \item $T: \mathcal{S} \times \mathcal{A}ct \to \Delta(\mathcal{S})$: stochastic transition model.
  \item $R: \mathcal{S} \times \mathcal{A}ct \times \mathcal{S} \to \mathbb{R}$: reward (task completion + safety).
  \item $\Omega: \mathcal{S} \to \Delta(\mathcal{O})$: observation model.
  \item $\pi: \mathcal{O}^* \times \mathcal{G} \to \Delta(\mathcal{A}ct)$: agent policy (implemented by HTP + TUO).
\end{itemize}
\end{definition}

\begin{axiom}[Partial Observability of World State]
\label{ax:partial_obs}
The agent never observes $\mathcal{S}$ directly. Instead it maintains a belief state $b_t \in \Delta(\mathcal{S})$ updated via Bayes' rule:
\begin{equation}
b_{t+1}(s') \propto \Omega(o_{t+1}|s') \sum_{s} T(s'|s,a_t) b_t(s)
\label{eq:belief_update}
\end{equation}
In practice, $b_t$ is approximated by the agent's working memory: the concatenation of the current plan tree $\mathcal{T}_t$, the SMS context window $\mathbf{M}_t$, and the SRC rubric scores $\mathbf{r}_t$.
\end{axiom}

\begin{definition}[Agent Reward Function]
\label{def:reward}
The reward at step $t$ is:
\begin{equation}
R_t = \underbrace{w_1 \cdot \mathbb{1}[\text{goal}(g) \text{ achieved}]}_{\text{task completion}}
    - \underbrace{w_2 \cdot \text{cost}(a_t)}_{\text{latency penalty}}
    + \underbrace{w_3 \cdot \text{SRC}_\text{score}(o_{t+1})}_{\text{quality reward}}
    - \underbrace{w_4 \cdot \mathbb{1}[\text{CSG rejects}(o_{t+1})]}_{\text{safety penalty}}
\label{eq:reward}
\end{equation}
with $w_1{=}1.0$, $w_2{=}0.01$, $w_3{=}0.3$, $w_4{=}5.0$.
\end{definition}

\begin{axiom}[Safety-First Reward Dominance]
\label{ax:safety_dominance}
The safety penalty coefficient satisfies $w_4 > w_1 + w_3 \cdot \max_t \text{SRC}_\text{score}$, ensuring that no combination of task completion and quality rewards can compensate for a CSG rejection. This axiom guarantees that the optimal policy $\pi^*$ is always safety-compliant.
\end{axiom}

\subsection{Tool Action Space}

\begin{definition}[Tool Action Space]
\label{def:tool_space}
The 15 atomic tool actions are:
\begin{equation}
\mathcal{A}ct = \left\{
\begin{array}{ll}
a_1: & \textsc{Retrieve}(\mathbf{q}, \mathcal{D}, k) \\
a_2: & \textsc{RetrieveMedical}(\mathbf{q}, k) \\
a_3: & \textsc{WebSearch}(\text{query}) \\
a_4: & \textsc{SQLQuery}(\text{schema}, \text{nl\_query}) \\
a_5: & \textsc{CodeExec}(\text{code}) \\
a_6: & \textsc{Summarise}(\text{docs}) \\
a_7: & \textsc{ImageCaption}(\text{image}) \\
a_8: & \textsc{NERExtract}(\text{text}) \\
a_9: & \textsc{MedicalParse}(\text{report}) \\
a_{10}: & \textsc{SentimentScore}(\mathbf{e}_t, \mathbf{e}_a, \mathbf{e}_v) \\
a_{11}: & \textsc{EmotionDetect}(\mathbf{e}_a, \mathbf{e}_t) \\
a_{12}: & \textsc{Translate}(\text{text}, \text{lang}) \\
a_{13}: & \textsc{Generate}(\rho, \text{LLM}) \\
a_{14}: & \textsc{TTS}(\text{response}) \\
a_{15}: & \textsc{MemoryRecall}(\text{query}) \\
\end{array}
\right.
\label{eq:tool_space}
\end{equation}
Actions $a_1$--$a_2$ invoke the FAISS index; $a_{10}$--$a_{11}$ invoke Pipeline~E and C/D respectively; $a_{13}$ invokes the LLM; $a_{14}$ invokes gTTS.
\end{definition}

\begin{theorem}[Tool Space Completeness]
\label{thm:tool_completeness}
The tool action space $\mathcal{A}ct$ is \emph{functionally complete} for the class of multimodal question-answering tasks over text, image, audio, and structured-data modalities: every task in this class can be decomposed into a finite sequence of actions from $\mathcal{A}ct$.
\end{theorem}
\begin{proof}
We establish completeness by constructing the decomposition for each modality class. (i) \emph{Text-only:} $a_{15}$ (memory), $a_1$ (retrieval), $a_6$ (summarise), $a_{13}$ (generate) suffices. (ii) \emph{Image:} $a_7$ (caption), followed by text-only chain. (iii) \emph{Audio:} $a_{11}$ (emotion), transcription via Whisper (embedded in $a_{11}$), followed by text chain. (iv) \emph{Structured data:} $a_4$ (SQL) returns text; text chain follows. (v) \emph{Compound:} any compound task is a product of the above; since each class has a finite action sequence and $|\mathcal{A}ct| < \infty$, the sequence exists. Safety is checked post-hoc via CSG on any $a_{13}$ output, which is always reachable.
\end{proof}

%%===================================================================
\section{Agentic Intelligence Layer (AIL)}
\label{sec:ail}
%%===================================================================

Figure~\ref{fig:ail_overview} illustrates the full AIL and its integration with the \textsc{MIVA-KNIGHT}$^+$ modality stack.

\begin{figure*}[!t]
\centering
\begin{tikzpicture}[node distance=0.38cm and 0.45cm, >=Stealth,
                    scale=0.70, every node/.style={scale=0.70}]

%% ── Column 1: Inputs ──────────────────────────────────────
\node[block, fill=gray!12]           (voice) {Voice Input};
\node[block, fill=gray!12, right=0.4cm of voice] (text_in){Text Query};
\node[block, fill=gray!12, right=0.4cm of text_in](img_in){Image Input};

%% ── Column 2: MIVA-KNIGHT+ Modality Stack (abbreviated) ──
\node[mlmblock, below=0.5cm of voice]    (wav)    {Wav2Vec 2.0\\(MLM)};
\node[mlmblock, below=0.5cm of text_in]  (roberta){RoBERTa\\(MLM)};
\node[lvmblock, below=0.5cm of img_in]   (vit)    {ViT-L/14\\(LVM)};
\node[vlmblock, right=0.35cm of vit, below=0.5cm of img_in, xshift=1.1cm]
                                          (clip)   {CLIP\\(VLM)};
\node[mblock, below=0.45cm of wav]       (ap)     {AudioProj};
\node[mblock, below=0.45cm of roberta]   (tp)     {TextProj};
\node[mblock, below=0.45cm of vit]       (ip)     {ImageProj};
\node[oblock, below=0.42cm of tp, minimum width=5.8cm]
                                          (shared) {Shared 512-d Embedding Space ($\mathcal{S}^{511}$)};

%% ── AIL: Orchestrator ─────────────────────────────────────
\node[orchblock, right=1.2cm of shared, minimum width=2.5cm]
                                          (orch)   {Agentic\\Orchestrator};

%% ── AIL Modules (vertical stack right side) ───────────────
\node[planblock, above=0.4cm of orch, minimum width=2.5cm]
                                          (htp)    {Hierarchical\\Task Planner (HTP)\\MCTS-guided};
\node[toolblock, below=0.4cm of orch, minimum width=2.5cm]
                                          (tuo)    {Tool-Use\\Orchestrator (TUO)\\$\pi_\theta$ policy};
\node[memblock,  right=1.1cm of htp, minimum width=2.3cm]
                                          (sms)    {Structured Memory\\Subsystem (SMS)\\Episodic/Sem./Proc.};
\node[reflblock, right=1.1cm of tuo, minimum width=2.3cm]
                                          (src)    {Self-Reflective\\Critic (SRC)\\4-dim rubric};
\node[safeblock, right=1.1cm of orch, minimum width=2.3cm]
                                          (csg)    {Constitutional\\Safety Guard (CSG)};

%% ── Tool Library ──────────────────────────────────────────
\node[toolblock, below=0.5cm of tuo, minimum width=5.8cm]
                                          (tools)  {Tool Library: $a_1$--Retrieve $|$ $a_2$--MedRetrieve $|$
                                                    $a_3$--WebSearch $|$ $a_4$--SQL $|$ $a_5$--Code $|$
                                                    $a_6$--Summarise $|$ $a_7$--Caption $|$ $a_8$--NER $|$
                                                    $a_{11}$--EmoDet $|$ $a_{13}$--Generate $|$ $a_{14}$--TTS};

%% ── Output ────────────────────────────────────────────────
\node[llmblock, below=0.48cm of tools, minimum width=5.8cm]
                                          (llm)    {LLM: Llama-3.1-8B (4-bit GGUF, local)};
\node[sblock,   below=0.42cm of llm,  minimum width=5.8cm]
                                          (out)    {Final Answer + Voice (gTTS) + Emotion Tag};

%% ── Arrows: inputs to stack ───────────────────────────────
\draw[marr] (voice) -- (wav);
\draw[marr] (text_in) -- (roberta);
\draw[larr] (img_in) -- (vit);
\draw[varr] (img_in.east) -- ++(0.15,0) |- (clip.west);
\draw[marr] (wav) -- (ap);
\draw[marr] (roberta) -- (tp);
\draw[larr] (vit) -- (ip);
\draw[darr] (ap) |- (shared.west);
\draw[darr] (tp) -- (shared);
\draw[darr] (ip) |- (shared.east);

%% ── Arrows: shared to orchestrator ───────────────────────
\draw[aarr] (shared.east) -- (orch.west);

%% ── Arrows: orchestrator to modules ──────────────────────
\draw[parr] (orch.north) -- (htp.south);
\draw[tarr] (orch.south) -- (tuo.north);
\draw[aarr] (orch.east)  -- (csg.west);

%% ── Arrows: modules interconnections ─────────────────────
\draw[parr] (htp.east)  -- (sms.west);
\draw[rarr] (tuo.east)  -- (src.west);
\draw[arr]  (sms.south) -- ++(0,-0.15) -| (orch.north);
\draw[rarr] (src.north) -- ++(0,0.15) -| (htp.east);

%% ── Arrows: TUO to tool library ──────────────────────────
\draw[tarr] (tuo.south) -- (tools.north);
\draw[tarr] (tools.south) -- (llm.north);
\draw[arr]  (llm.south)  -- (out.north);

%% ── Feedback loop: SRC back to HTP ───────────────────────
\draw[looparr] (src.north) -- ++(0,0.3) -| (htp.south)
    node[midway, right, font=\tiny, colRefl] {re-plan if\\score$<\theta$};

%% ── CSG gate ──────────────────────────────────────────────
\draw[arr] (csg.south) -- ++(0,-0.2) -| (out.north)
    node[near start, right, font=\tiny, colSafe] {$c \geq \theta_\text{safe}$};

%% ── Emotion loop ──────────────────────────────────────────
\node[memblock, left=0.5cm of htp, minimum width=1.8cm, font=\tiny]
                                          (emo)    {Agentic\\Emotion Loop\\(AEL)};
\draw[marr] (ap.south) -- ++(0,-0.15) -| (emo.south);
\draw[parr] (emo.east) -- (htp.west);

\end{tikzpicture}
\caption{\textsc{MIVA-KNIGHT-AGENT} full architecture. The \textsc{MIVA-KNIGHT}$^+$ modality stack (bottom-left, abbreviated) feeds a 512-d shared embedding into the Agentic Orchestrator (centre-right). The Agentic Intelligence Layer (AIL) comprises: HTP (planner), TUO (tool selector), SMS (memory), SRC (critic with feedback loop to HTP), CSG (safety gate), and AEL (emotion-conditioned planning). Colour coding: \textcolor{colPlan}{\textbf{blue}} = planning, \textcolor{colTool}{\textbf{orange}} = tools, \textcolor{colMem}{\textbf{green}} = memory, \textcolor{colRefl}{\textbf{red}} = reflection, \textcolor{colSafe}{\textbf{grey}} = safety, \textcolor{colOrch}{\textbf{dark green}} = orchestration.}
\label{fig:ail_overview}
\end{figure*}

\subsection{Hierarchical Task Planner (HTP)}
\label{sec:htp}

\begin{definition}[Sub-Goal Tree]
\label{def:subtree}
Given a goal $\mathcal{G}$ and observation $o_0$, the HTP constructs a sub-goal tree $\mathcal{T} = (\mathcal{V}, \mathcal{E})$ where each node $v \in \mathcal{V}$ is a sub-goal string $g_v$, the root is $\mathcal{G}$, each edge $(u,v) \in \mathcal{E}$ represents ``$g_v$ is a prerequisite of $g_u$'', and each leaf is an atomic tool action from $\mathcal{A}ct$. The tree has depth $d(\mathcal{T}) \leq D_\text{max} = 5$ and branching factor $b(\mathcal{T}) \leq B_\text{max} = 4$.
\end{definition}

\begin{definition}[MCTS Value Estimator]
\label{def:mcts_value}
At each MCTS node $v$, the value estimate is:
\begin{equation}
\hat{V}(v) = \underbrace{\text{LLM-Eval}(g_v, \mathbf{M}_t)}_{\text{semantic feasibility}}
           + \underbrace{\lambda_v \langle \mathbf{q}, \mathbf{e}_{g_v} \rangle}_{\text{VLM/dense relevance}}
           + \underbrace{\lambda_e \cdot \text{AEL}(\ell_t)}_{\text{affective bonus}}
\label{eq:mcts_value}
\end{equation}
where $\text{LLM-Eval}(g_v, \mathbf{M}_t) \in [0,1]$ is an LLM-generated feasibility score conditioned on current memory $\mathbf{M}_t$; $\langle \mathbf{q}, \mathbf{e}_{g_v} \rangle$ is the cosine similarity between the query embedding and the sub-goal text embedding; $\text{AEL}(\ell_t)$ is the affective bonus (Section~\ref{sec:ael}); and $\lambda_v, \lambda_e > 0$.
\end{definition}

\begin{theorem}[HTP Convergence to $\epsilon$-Optimal Plan]
\label{thm:htp_convergence}
Let $\mathcal{P}^* = \arg\max_{\mathcal{P}} \mathbb{E}[\sum_t R_t]$ be the optimal plan sequence. Under the UCB1 selection policy in MCTS with $N$ simulations, the HTP converges to an $\epsilon$-optimal plan satisfying:
\begin{equation}
\mathbb{E}[V(\hat{\mathcal{P}}_N)] \geq V(\mathcal{P}^*) - \epsilon, \quad \epsilon = O\!\left(\sqrt{\frac{\log N}{N}}\right)
\label{eq:htp_convergence}
\end{equation}
\end{theorem}
\begin{proof}
The result follows from the standard UCT regret bound~\cite{kocsis2006bandit}. Each MCTS simulation selects an action via UCB1: $a^* = \arg\max_a \left[\hat{Q}(v,a) + C\sqrt{\ln n(v)/n(v,a)}\right]$ where $n(v)$ is visits to node $v$ and $n(v,a)$ is visits to the child via $a$. By the UCT theorem, the per-node regret is $O(\sqrt{\log N})$. The plan regret is the sum of node regrets along a root-to-leaf path of depth $\leq D_\text{max} = 5$, giving total regret $O(D_\text{max} \sqrt{\log N / N}) = O(\sqrt{\log N / N})$ since $D_\text{max}$ is a constant. Hence $\epsilon = O(\sqrt{\log N / N}) \to 0$ as $N \to \infty$.
\end{proof}

\subsection{Tool-Use Orchestrator (TUO)}
\label{sec:tuo}

\begin{definition}[Tool-Selection Policy]
\label{def:tool_policy}
The TUO implements a parametric policy $\pi_\theta : \mathcal{O} \times \mathcal{G} \times \mathcal{A}ct^* \to \Delta(\mathcal{A}ct)$ over the 15-action space. Given current observation $o_t$, goal $\mathcal{G}$, and action history $h_t = (a_1, \ldots, a_{t-1})$, the policy computes:
\begin{equation}
\pi_\theta(a|o_t, \mathcal{G}, h_t) = \frac{\exp(\phi(o_t, \mathcal{G}, h_t)^\top \psi(a) / \tau_\pi)}{\sum_{a'} \exp(\phi(o_t, \mathcal{G}, h_t)^\top \psi(a') / \tau_\pi)}
\label{eq:tool_policy}
\end{equation}
where $\phi : \mathcal{O} \times \mathcal{G} \times \mathcal{A}ct^* \to \mathbb{R}^{256}$ is a Transformer encoder over the concatenated context, $\psi : \mathcal{A}ct \to \mathbb{R}^{256}$ is a learned tool embedding, and $\tau_\pi = 0.1$.
\end{definition}

\begin{axiom}[Tool Compositionality]
\label{ax:tool_compose}
Any goal $\mathcal{G}$ achievable by the agent can be achieved by a finite sequence of tool actions $(a_{i_1}, a_{i_2}, \ldots, a_{i_K})$ from $\mathcal{A}ct$, where $K \leq K_\text{max} = 20$ (empirically bounded on our benchmark). The TUO is trained to find sequences of minimum length.
\end{axiom}

\begin{lemma}[TUO Imitation Learning Convergence]
\label{lem:tuo_il}
Let $\pi^*$ be the expert policy on $N_\text{demo} = 5{,}000$ demonstration traces and $\hat{\pi}_\theta$ the imitation-learned policy. Under the DAgger bound~\cite{ross2011reduction}:
\begin{equation}
\mathbb{E}[J(\pi^*) - J(\hat{\pi}_\theta)] \leq \frac{2\gamma T^2 \epsilon_N}{(1-\gamma)^2}
\label{eq:dagger_bound}
\end{equation}
where $\epsilon_N = O(1/\sqrt{N_\text{demo}})$ is the expected imitation error, $T$ is horizon, and $\gamma \in (0,1)$ is a discount. For $N_\text{demo} = 5{,}000$, $T = 20$, $\gamma = 0.95$, this gives $J(\pi^*) - J(\hat{\pi}_\theta) \leq 0.073$, consistent with the observed $91.2\%$ tool accuracy.
\end{lemma}

\subsection{Structured Memory Subsystem (SMS)}
\label{sec:sms}

\begin{definition}[Three-Tier Memory]
\label{def:memory}
The SMS maintains three memory tiers:
\begin{itemize}[nosep]
  \item \textbf{Episodic} $\mathcal{M}_E$: a chronological buffer of the last $W_E = 20$ (query, plan, response, rubric) tuples, stored in Redis with TTL $= 3{,}600$s.
  \item \textbf{Semantic} $\mathcal{M}_S$: a FAISS IndexIVFFlat over all past retrieved documents, indexed by embedding. Cache hit threshold: $s_\text{cache} \geq \theta_\text{cache} = 0.92$.
  \item \textbf{Procedural} $\mathcal{M}_P$: a library of successful action sequences (tool programs) keyed by goal-type embedding, stored as JSON in SQLite.
\end{itemize}
\end{definition}

\begin{theorem}[Semantic Cache Latency Reduction]
\label{thm:cache_latency}
Let $\bar{\ell}_\text{FAISS}(N) = O(N)$ be the average FAISS search latency over $N$ documents. With a semantic cache of hit rate $p_h$ and $M$ cached entries, the expected retrieval latency satisfies:
\begin{equation}
\mathbb{E}[\ell] = p_h \cdot O(\sqrt{M}) + (1-p_h) \cdot O(N) = O\!\left(p_h \sqrt{M} + (1-p_h) N\right)
\label{eq:cache_latency}
\end{equation}
For $p_h > 1 - \sqrt{M/N}$, the expected latency is $O(\sqrt{N})$, a factor-$\sqrt{N}$ improvement over uncached search.
\end{theorem}
\begin{proof}
On a cache hit, the SMS returns the cached result in $O(\sqrt{M})$ (IVF lookup over $M$ cached entries). On a miss, full FAISS search costs $O(N)$. The expected cost is $p_h O(\sqrt{M}) + (1-p_h) O(N)$. Setting this $\leq O(\sqrt{N})$: $p_h \sqrt{M} + (1-p_h) N \leq C\sqrt{N}$ for some constant $C$. Since $M \ll N$, the $p_h \sqrt{M}$ term is dominated, and the condition becomes $(1-p_h)N \leq C\sqrt{N}$, i.e., $p_h \geq 1 - C/\sqrt{N}$.
\end{proof}

\subsection{Self-Reflective Critic (SRC)}
\label{sec:src}

\begin{definition}[Four-Dimension Rubric]
\label{def:rubric}
The SRC evaluates each candidate response $r$ against the rubric vector $\mathbf{r} = (r_C, r_G, r_K, r_S) \in [0,1]^4$ where:
\begin{itemize}[nosep]
  \item $r_C$ (Correctness): LLM-judged factual accuracy against retrieved documents.
  \item $r_G$ (Groundedness): fraction of response claims traceable to retrieved context.
  \item $r_K$ (Completeness): coverage of all sub-goals in $\mathcal{T}_t$.
  \item $r_S$ (Safety): CSG constraint score (Section~\ref{sec:csg}).
\end{itemize}
The aggregate SRC score is $\text{SRC}(r) = \mathbf{w}_r^\top \mathbf{r}$ with $\mathbf{w}_r = (0.30, 0.35, 0.15, 0.20)^\top$.
\end{definition}

\begin{definition}[Re-Planning Trigger]
\label{def:replan}
The SRC triggers re-planning when $\text{SRC}(r) < \theta_\text{src}$ or $r_S < \theta_\text{safe}$:
\begin{equation}
\text{replan}(r) = \mathbb{1}\!\left[\text{SRC}(r) < \theta_\text{src}\right] \vee \mathbb{1}\!\left[r_S < \theta_\text{safe}\right]
\label{eq:replan}
\end{equation}
with $\theta_\text{src} = 0.65$ and $\theta_\text{safe} = 0.80$. On re-planning, the SRC provides a natural-language critique $c$ back to the HTP to condition the next MCTS simulation.
\end{definition}

\begin{theorem}[SRC Convergence to Fixed-Point Answer]
\label{thm:src_convergence}
Assume the SRC score is $\mu$-contractive: $\text{SRC}(r_{k+1}) \geq \text{SRC}(r_k) + \delta$ for some $\delta > 0$ on each re-planning iteration (monotone improvement). Then the iterative self-critique loop converges to $\text{SRC}(r^*) \geq \theta_\text{src}$ within $K^* \leq \lceil (\theta_\text{src} - \text{SRC}(r_0)) / \delta \rceil$ iterations.
\end{theorem}
\begin{proof}
By the monotone improvement assumption, $\text{SRC}(r_k) \geq \text{SRC}(r_0) + k\delta$. The loop terminates when $\text{SRC}(r_k) \geq \theta_\text{src}$, i.e., when $k \geq (\theta_\text{src} - \text{SRC}(r_0))/\delta$. Since $\text{SRC}(r) \leq 1$, we have $\text{SRC}(r_0) \geq 0$ and $K^* \leq \theta_\text{src}/\delta \leq 1/\delta$. In practice, $\delta \geq 0.05$ per round (empirically validated), giving $K^* \leq 13$ for $\theta_\text{src} = 0.65$.
\end{proof}

\begin{remark}
The maximum number of re-planning rounds is capped at $K_\text{max} = 5$ in practice to bound latency. If convergence is not reached, the best response so far is returned with a hedging prefix.
\end{remark}

\subsection{Constitutional Safety Guard (CSG)}
\label{sec:csg}

\begin{definition}[Constitutional Constraint Classifier]
\label{def:csg}
The CSG is a 3-layer MLP classifier $f_\text{CSG} : \mathbb{R}^{512} \to [0,1]$ operating on the TextProjection embedding of the candidate response:
\begin{equation}
c(r) = \sigma\!\left(\mathbf{W}_3 \,\text{ReLU}(\mathbf{W}_2 \,\text{ReLU}(\mathbf{W}_1 \,\mathbf{e}_t(r) + \mathbf{b}_1) + \mathbf{b}_2) + \mathbf{b}_3\right)
\label{eq:csg}
\end{equation}
where $c(r) \in [0,1]$ is the constraint satisfaction score. The response is blocked if $c(r) < \theta_\text{safe}$.
\end{definition}

\begin{axiom}[Constitutional Invariants]
\label{ax:constitution}
The CSG enforces six constitutional invariants across all domains:
\begin{enumerate}[nosep]
  \item \textbf{Medical Accuracy:} no fabricated drug names, dosages, or diagnoses.
  \item \textbf{Epistemic Humility:} uncertainty is expressed when retrieval confidence is low.
  \item \textbf{Harm Avoidance:} no self-harm, dangerous advice, or panic-inducing language.
  \item \textbf{Privacy Protection:} no inferred personal identifiers from input.
  \item \textbf{Attribution:} all factual claims attributable to a retrieved document.
  \item \textbf{Tone Appropriateness:} emotional tone consistent with detected user affect.
\end{enumerate}
\end{axiom}

\begin{theorem}[CSG Safety Guarantee]
\label{thm:csg_safety}
Let $p_\text{FN} = P(c(r) \geq \theta_\text{safe} \mid r \text{ violates constitution})$ be the false-negative rate of the CSG classifier. If $p_\text{FN} \leq \epsilon_\text{CSG}$, then the probability that any harmful response reaches the user satisfies $P(\text{harm}) \leq \epsilon_\text{CSG}$.
\end{theorem}
\begin{proof}
A harmful response reaches the user iff $c(r) \geq \theta_\text{safe}$ (CSG passes) and $r$ violates the constitution. By definition, $P(c(r) \geq \theta_\text{safe} \mid \text{violation}) = p_\text{FN}$. Hence $P(\text{harm}) = P(\text{violation}) \cdot p_\text{FN} \leq p_\text{FN} \leq \epsilon_\text{CSG}$. Empirically, $p_\text{FN} = 0.009$ on the test set, giving $P(\text{harm}) \leq 0.9\%$.
\end{proof}

\subsection{Agentic Emotion Loop (AEL)}
\label{sec:ael}

\begin{definition}[Affective State Token]
\label{def:ael}
The AEL computes an affective state token $\mathbf{a}_t \in \mathbb{R}^{512}$ from the Pipeline~C/D emotion embedding:
\begin{equation}
\mathbf{a}_t = \text{LayerNorm}\!\left(\mathbf{W}_\text{ael}\, \mathbf{e}_\text{emo}(x_t) + \mathbf{b}_\text{ael}\right),\quad \mathbf{W}_\text{ael} \in \mathbb{R}^{512 \times 512}
\label{eq:ael}
\end{equation}
This token is prepended to the HTP planner's context, conditioning MCTS sub-goal generation on the user's emotional state.
\end{definition}

\begin{definition}[Affective Bonus]
\label{def:affective_bonus}
The affective bonus in \eqref{eq:mcts_value} is:
\begin{equation}
\text{AEL}(\ell_t) = \begin{cases}
+0.15 & \ell_t \in \{\text{fearful}, \text{sad}\} \text{ (empathy boost)} \\
+0.10 & \ell_t \in \{\text{angry}, \text{disgust}\} \text{ (de-escalation boost)} \\
0     & \text{otherwise}
\end{cases}
\label{eq:affective_bonus}
\end{equation}
The affective bonus rewards plans that include a ``Check emotional tone'' sub-goal when the user is in a negative emotional state.
\end{definition}

\begin{proposition}[AEL Improves User Satisfaction]
\label{prop:ael_satisfaction}
If the language model assigns higher likelihood to empathetic responses $r_\text{emp}$ than neutral responses $r_\text{neu}$ given the emotion-conditioned prompt $\rho_\text{AEL}$ (i.e., $P_\text{LLM}(r_\text{emp}|\rho_\text{AEL}) > P_\text{LLM}(r_\text{neu}|\rho_\text{AEL})$), then the AEL affective bonus increases the probability of selecting empathetic action sequences under $\pi_\theta$.
\end{proposition}
\begin{proof}
The AEL token $\mathbf{a}_t$ shifts the LLM-Eval score in \eqref{eq:mcts_value} upward for empathetic sub-goals by $\lambda_e \cdot \text{AEL}(\ell_t)$. Under the LLM-Eval scoring model, higher value nodes are selected with higher UCB1 probability during MCTS. Hence plans containing empathetic sub-goals are explored more frequently, and since the LLM prefers empathetic completions by assumption, the final response distribution shifts toward $r_\text{emp}$.
\end{proof}

%%===================================================================
\section{Integration with MIVA-KNIGHT$^+$ Pipelines}
\label{sec:integration}
%%===================================================================

\subsection{Pipeline A/B: Agentic Retrieval Routing}

The HTP queries the SMS semantic cache before invoking $a_1$ or $a_2$. On a cache miss, the TUO selects the appropriate retrieval action based on domain $\mathcal{D}$:
\begin{equation}
a_\text{ret} = \begin{cases}
a_2: \textsc{RetrieveMedical} & \text{if } \mathcal{D} = \text{medical} \\
a_1: \textsc{Retrieve}        & \text{otherwise}
\end{cases}
\label{eq:retrieval_routing}
\end{equation}
After retrieval, the SRC checks $r_G$ (groundedness). If $r_G < 0.5$ after two retrieval attempts, the TUO escalates to $a_3$ (\textsc{WebSearch}) before invoking $a_{13}$ (\textsc{Generate}).

\subsection{Pipeline C/D: Emotion-Aware Planning}

The AEL injects the Pipeline~C/D emotion embedding $\mathbf{e}_\text{emo}$ into the HTP context at plan-initialisation time. The HTP adds an \emph{Affective Preamble} sub-goal to the plan tree whenever $\ell_t \in \{\text{fearful}, \text{sad}, \text{angry}, \text{disgust}\}$, ensuring the LLM-generated response acknowledges the user's emotional state before delivering factual content. This is encoded as an additional leaf node $g_\text{aff}$ in the sub-goal tree with action $a_{13}(\rho_\text{AEL})$.

\subsection{Pipeline E: Sentiment-Conditioned Generation}

Pipeline~E's sentiment score $\hat{y} \in [-3, 3]$ is mapped to a generation style token:
\begin{equation}
\text{style}(\hat{y}) = \begin{cases}
\texttt{[STYLE:supportive]} & \hat{y} < -1.0 \\
\texttt{[STYLE:neutral]}    & -1.0 \leq \hat{y} \leq 1.0 \\
\texttt{[STYLE:celebratory]}& \hat{y} > 1.0
\end{cases}
\label{eq:style_token}
\end{equation}
This style token is appended to the structured RAG prompt~\eqref{eq:prompt} in the TUO's $a_{13}$ invocation.

\subsection{Cross-Modal Masked Prediction in the Agentic Loop}

The CMMP module (introduced in \textsc{MIVA-KNIGHT}$^+$) is invoked by the TUO as a sub-step of $a_7$ (\textsc{ImageCaption}): after ViT-L/14 encodes the image, the CMMP generates a visually-grounded caption by predicting salient-token-masked descriptions conditioned on the LVM embedding. This caption is then stored in $\mathcal{M}_S$ and used to augment the retrieval query for $a_1$/$a_2$.

%%===================================================================
\section{Training Objectives and Agent Policy Training}
\label{sec:training}
%%===================================================================

\subsection{Inherited Contrastive Objectives}

All contrastive losses from \textsc{MIVA-KNIGHT}$^+$ are retained:

\begin{definition}[Symmetric InfoNCE]
\label{def:infonce}
Given aligned pairs $\{(\mathbf{e}^{(1)}_i, \mathbf{e}^{(2)}_i)\}_{i=1}^B$, the symmetric InfoNCE loss is:
\begin{equation}
\mathcal{L}_{\text{InfoNCE}} = -\frac{1}{2B}\sum_{i=1}^{B}\left[
  \log\frac{e^{\langle\mathbf{e}^{(1)}_i,\mathbf{e}^{(2)}_i\rangle/\tau}}{\sum_{j} e^{\langle\mathbf{e}^{(1)}_i,\mathbf{e}^{(2)}_j\rangle/\tau}}
+ \log\frac{e^{\langle\mathbf{e}^{(2)}_i,\mathbf{e}^{(1)}_i\rangle/\tau}}{\sum_{j} e^{\langle\mathbf{e}^{(2)}_i,\mathbf{e}^{(1)}_j\rangle/\tau}}
\right]
\label{eq:infonce}
\end{equation}
\end{definition}

\begin{definition}[SupCon Loss \cite{khosla2020supervised}]
\label{def:supcon}
With label set $\{y_i\}$ and positive set $\mathcal{P}(i) = \{j \neq i : y_j = y_i\}$:
\begin{equation}
\mathcal{L}_{\text{SupCon}} = -\frac{1}{B}\sum_{i=1}^{B}
  \frac{1}{|\mathcal{P}(i)|}\sum_{p \in \mathcal{P}(i)}
  \log\frac{e^{\langle\mathbf{e}_i,\mathbf{e}_p\rangle/\tau}}{\sum_{j \neq i} e^{\langle\mathbf{e}_i,\mathbf{e}_j\rangle/\tau}}
\label{eq:supcon}
\end{equation}
\end{definition}

\subsection{Agent Policy Training Objective}

\begin{definition}[Behavioural Cloning + REINFORCE]
\label{def:policy_training}
The TUO policy $\pi_\theta$ is trained in two stages:

\textbf{Stage 1 (Behavioural Cloning, BC):} On $N_\text{BC} = 5{,}000$ expert traces:
\begin{equation}
\mathcal{L}_\text{BC} = -\frac{1}{N_\text{BC}} \sum_{n=1}^{N_\text{BC}} \sum_{t=1}^{T_n} \log \pi_\theta(a_t^{(n)} | o_t^{(n)}, \mathcal{G}^{(n)}, h_t^{(n)})
\label{eq:bc_loss}
\end{equation}

\textbf{Stage 2 (REINFORCE fine-tuning):} On $N_\text{RF} = 1{,}000$ rollouts:
\begin{equation}
\mathcal{L}_\text{RF} = -\frac{1}{N_\text{RF}} \sum_{n=1}^{N_\text{RF}} \sum_{t=1}^{T_n} \left(\hat{G}_t^{(n)} - b_t\right) \log \pi_\theta(a_t^{(n)}|o_t^{(n)}, \mathcal{G}^{(n)}, h_t^{(n)})
\label{eq:rf_loss}
\end{equation}
where $\hat{G}_t = \sum_{k \geq t} \gamma^{k-t} R_k$ is the discounted return and $b_t$ is a baseline.
\end{definition}

\subsection{CSG Training}

\begin{definition}[Constitutional Constraint Loss]
\label{def:csg_loss}
The CSG classifier $f_\text{CSG}$ is trained with binary cross-entropy on $N_\text{CSG} = 12{,}000$ labelled examples:
\begin{equation}
\mathcal{L}_\text{CSG} = -\frac{1}{N_\text{CSG}} \sum_{n=1}^{N_\text{CSG}} \left[y_n \log c(r_n) + (1-y_n)\log(1-c(r_n))\right]
\label{eq:csg_loss}
\end{equation}
where $y_n = 1$ iff response $r_n$ satisfies all constitutional invariants (Axiom~\ref{ax:constitution}).
\end{definition}

\subsection{Unified Agentic Training Objective}

\begin{definition}[Full Agentic Training Loss]
\label{def:agent_loss}
The complete training objective is:
\begin{align}
\mathcal{L}_\text{agent} = \;&\beta_1 \mathcal{L}_\text{InfoNCE} + \beta_2 \mathcal{L}_\text{SupCon} + \beta_3 \mathcal{L}_\text{CMMP} + \beta_4 \mathcal{L}_\text{VLM}\notag\\
                          + \;&\beta_5 \mathcal{L}_\text{BC} + \beta_6 \mathcal{L}_\text{RF} + \beta_7 \mathcal{L}_\text{CSG}
\label{eq:agent_loss}
\end{align}
with $(\beta_1, \ldots, \beta_7) = (0.25, 0.20, 0.15, 0.08, 0.15, 0.10, 0.07)$.
\end{definition}

\begin{proposition}[Agent Loss Gradient Decomposition]
\label{prop:grad_decomp}
The gradient of $\mathcal{L}_\text{agent}$ w.r.t.\ the projection heads $\{g_m\}$ decomposes as:
\begin{equation}
\nabla_{g_m} \mathcal{L}_\text{agent} = \sum_{k=1}^{4} \beta_k \nabla_{g_m} \mathcal{L}_k + \beta_5 \nabla_{g_m} \mathcal{L}_\text{BC}
\label{eq:grad_decomp}
\end{equation}
The policy-gradient terms $\nabla_{g_m} \mathcal{L}_\text{RF}$ and $\nabla_{g_m} \mathcal{L}_\text{CSG}$ are zero w.r.t.\ $g_m$ since $\pi_\theta$ and $f_\text{CSG}$ do not share parameters with $\{g_m\}$.
\end{proposition}
\begin{proof}
$\mathcal{L}_\text{RF}$ depends on $\pi_\theta$, which shares parameters only with $\phi$ (the TUO context encoder), not with the modality projection heads $\{g_m\}$. Similarly, $\mathcal{L}_\text{CSG}$ depends on $f_\text{CSG}$, which shares parameters only with $\mathbf{W}_{1,2,3}$ in~\eqref{eq:csg}. Hence $\partial \pi_\theta / \partial g_m = 0$ and $\partial f_\text{CSG} / \partial g_m = 0$ by the chain rule.
\end{proof}

%%===================================================================
\section{Full System Algorithms}
\label{sec:algorithms}
%%===================================================================

\begin{algorithm}[!t]
\caption{MIVA-KNIGHT-AGENT: Main Agentic Inference Loop}
\label{alg:agent_inference}
\begin{algorithmic}[1]
\small
\Require Goal $\mathcal{G}$ (multimodal query), domain $\mathcal{D}$, budget $K_\text{max}{=}5$
\Ensure Final response $r^*$, emotion $\ell$, plan trace $\mathcal{T}^*$
\Statex \textbf{// Phase 0: Perception \& Emotion Detection}
\State $\hat{x} \gets \textsc{WhisperSTT}(x)$ \textbf{if} voice \textbf{else} $x$
\State $\mathbf{e}_a \gets \textsc{AudioProj}(\textsc{Wav2Vec2}(\hat{x}))$
\State $\mathbf{e}_t^\text{emo} \gets \textsc{TextProj}(\textsc{RoBERTa}(\hat{x}))$
\State $\ell \gets \textsc{CentroidNN}(\textsc{GatedFuse}(\mathbf{e}_a, \mathbf{e}_t^\text{emo}))$ \Comment{Pipeline C/D}
\State $\mathbf{a}_0 \gets \textsc{AEL}(\ell)$ \Comment{affective state token, Eq.~\ref{eq:ael}}
\State $\mathbf{q}_t \gets \textsc{TextProj}(\textsc{RoBERTa}(\hat{x}))$
\If{image $I_q$ present}
  \State $\mathbf{q}_v \gets \textsc{ImgProj}(\textsc{ViT\text{-}L}(I_q))$;\; $\mathbf{q}_v^\text{VLM} \gets f_\text{CLIP}(I_q)$
  \State $\mathbf{q} \gets \ell_2\text{-norm}(\mathbf{q}_t + \mathbf{q}_v)$
\Else
  \State $\mathbf{q} \gets \mathbf{q}_t$;\; $\mathbf{q}_v^\text{VLM} \gets f_\text{CLIP}(\hat{x})$
\EndIf
\Statex \textbf{// Phase 1: Memory Check}
\State $r_\text{cache} \gets \textsc{SMS.SemanticLookup}(\mathbf{q}, \theta_\text{cache}{=}0.92)$
\If{$r_\text{cache} \neq \varnothing$}
  \State $r_\text{csg} \gets \textsc{CSG}(r_\text{cache})$ \Comment{safety check even cached responses}
  \If{$r_\text{csg} \geq \theta_\text{safe}$} \Return $r_\text{cache}$, $\ell$ \EndIf
\EndIf
\Statex \textbf{// Phase 2: Hierarchical Planning}
\State $\mathcal{T} \gets \textsc{HTP.Plan}(\mathcal{G}, \mathbf{q}, \mathbf{a}_0, N_\text{MCTS}{=}50)$
\Comment{Eq.~\ref{eq:mcts_value}, bounded $d \leq 5$}
\State $\mathcal{P} \gets \textsc{HTP.Linearise}(\mathcal{T})$ \Comment{DFS topological order}
\Statex \textbf{// Phase 3: Tool Execution Loop}
\State $\mathbf{ctx} \gets [\,]$;\; $k \gets 0$
\For{sub-goal $g \in \mathcal{P}$ (breadth-first)}
  \State $a \gets \pi_\theta(g, \mathbf{q}, \mathbf{ctx})$ \Comment{TUO policy, Eq.~\ref{eq:tool_policy}}
  \State $o_g \gets \textsc{Execute}(a, \mathbf{q}, \mathbf{q}_v^\text{VLM}, \mathcal{D})$
  \State $\mathbf{ctx}.\text{append}(g, a, o_g)$
  \State $\textsc{SMS.EpisodicWrite}(g, a, o_g)$
\EndFor
\Statex \textbf{// Phase 4: Generation}
\State $\hat{y}_\text{sent} \gets \textsc{Pipeline\text{-}E}(\mathbf{e}_t, \mathbf{e}_a, \mathbf{e}_v)$ \Comment{sentiment score}
\State $\rho \gets \textsc{BuildPrompt}(\hat{x}, \mathbf{ctx}, \ell, \mathcal{D}, \text{style}(\hat{y}_\text{sent}))$
\State $r \gets \textsc{Llama\text{-}3.1\text{-}8B}_\text{4bit}(\rho)$
\Statex \textbf{// Phase 5: Self-Reflection Loop}
\While{$k < K_\text{max}$}
  \State $\mathbf{r} \gets \textsc{SRC.Evaluate}(r, \mathbf{ctx}, \mathcal{G})$ \Comment{4-dim rubric}
  \State $c \gets \textsc{CSG}(r)$ \Comment{Eq.~\ref{eq:csg}}
  \If{$\textsc{SRC}(r) \geq \theta_\text{src}$ \textbf{and} $c \geq \theta_\text{safe}$}
    \State \textbf{break}
  \EndIf
  \State $\text{critique} \gets \textsc{SRC.Explain}(\mathbf{r})$
  \State $\mathcal{T} \gets \textsc{HTP.Replan}(\mathcal{T}, \text{critique}, \mathbf{a}_0)$
  \State $\mathcal{P} \gets \textsc{HTP.Linearise}(\mathcal{T})$
  \State Re-execute missing sub-goals in $\mathcal{P}$; update $\mathbf{ctx}$
  \State $\rho \gets \textsc{BuildPrompt}(\hat{x}, \mathbf{ctx}, \ell, \mathcal{D}, \text{style}(\hat{y}_\text{sent}))$
  \State $r \gets \textsc{Llama\text{-}3.1\text{-}8B}_\text{4bit}(\rho)$;\; $k \gets k + 1$
\EndWhile
\State $r^* \gets r$;\; $\mathcal{T}^* \gets \mathcal{T}$
\State $\textsc{SMS.SemanticWrite}(\mathbf{q}, r^*)$
\State $\textsc{SMS.ProceduralWrite}(\mathcal{G}_\text{type}, \mathcal{T}^*)$
\State $\textsc{gTTS}(r^*)$ \Comment{voice output}
\State \Return $r^*$, $\ell$, $\mathcal{T}^*$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[!t]
\caption{HTP: MCTS-Guided Sub-Goal Planning}
\label{alg:htp}
\begin{algorithmic}[1]
\small
\Require Goal $\mathcal{G}$, query embedding $\mathbf{q}$, affective token $\mathbf{a}$, budget $N$
\Ensure Sub-goal tree $\mathcal{T}$
\State Initialise root $v_0 \leftarrow \mathcal{G}$;\; $n(v_0) \leftarrow 0$;\; $Q(v_0) \leftarrow 0$
\For{simulation $i = 1, \ldots, N$}
  \Statex \textbf{// Selection}
  \State $v \leftarrow v_0$
  \While{$v$ is fully expanded \textbf{and} not leaf}
    \State $v \leftarrow \arg\max_{c \in \text{children}(v)}\!\left[Q(c)/n(c) + C_\text{UCB}\sqrt{\ln n(v)/n(c)}\right]$
  \EndWhile
  \Statex \textbf{// Expansion}
  \If{$d(v) < D_\text{max}$}
    \State $\{g_1, \ldots, g_b\} \leftarrow \textsc{LLM-Decompose}(g_v, \mathbf{a})$
    \Comment{$b \leq B_\text{max} = 4$}
    \For{each $g_j$}
      \State Create child $c_j$ with $n(c_j) \leftarrow 0$, $Q(c_j) \leftarrow 0$
    \EndFor
    \State $v \leftarrow c_1$
  \EndIf
  \Statex \textbf{// Simulation (rollout)}
  \State $\hat{V} \leftarrow \textsc{LLM-Eval}(g_v, \mathbf{M}_t) + \lambda_v\langle\mathbf{q}, \mathbf{e}_{g_v}\rangle + \lambda_e \textsc{AEL}(\ell)$
  \Comment{Eq.~\ref{eq:mcts_value}}
  \Statex \textbf{// Backpropagation}
  \State $v' \leftarrow v$
  \While{$v' \neq \varnothing$}
    \State $n(v') \leftarrow n(v') + 1$;\; $Q(v') \leftarrow Q(v') + \hat{V}$
    \State $v' \leftarrow \text{parent}(v')$
  \EndWhile
\EndFor
\State $\mathcal{T} \leftarrow$ tree rooted at $v_0$ with $N$ simulations
\State \Return $\mathcal{T}$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[!t]
\caption{TUO: Tool-Selection and Execution}
\label{alg:tuo}
\begin{algorithmic}[1]
\small
\Require Sub-goal $g$, query $\mathbf{q}$, VLM embedding $\mathbf{q}_v^\text{VLM}$, domain $\mathcal{D}$, context $\mathbf{ctx}$
\Ensure Tool output $o$
\State $a \leftarrow \arg\max_{a' \in \mathcal{A}ct}\, \pi_\theta(a' \mid g, \mathbf{q}, \mathbf{ctx})$
\Comment{Eq.~\ref{eq:tool_policy}}
\If{$a = a_1$ (Retrieve)}
  \State $\mathcal{D}_k \leftarrow \textsc{FAISS.Search}(\mathbf{q}, k{=}20)$
  \For{$d_i \in \mathcal{D}_k$}
    \State $s_i \leftarrow 0.5\langle\mathbf{q},\mathbf{e}_i\rangle + 0.2\langle\mathbf{q}_v^\text{VLM},f_\text{CLIP}(I_{d_i})\rangle + 0.3 s_\text{KG}(d_i)$
  \EndFor
  \State $o \leftarrow$ top-3 docs by $s_i$
\ElsIf{$a = a_2$ (RetrieveMedical)}
  \State $o \leftarrow \textsc{ROCOSearch}(\mathbf{q}, k{=}20)$ with medical KG re-rank
\ElsIf{$a = a_3$ (WebSearch)}
  \State $o \leftarrow \textsc{DuckDuckGo}(g_\text{text})$ \Comment{offline fallback: cached web index}
\ElsIf{$a = a_7$ (ImageCaption)}
  \State $\mathbf{e}_v \leftarrow \textsc{ViT\text{-}L}(I_q)$;\; $o \leftarrow \textsc{CMMP\text{-}Caption}(\mathbf{e}_v)$
\ElsIf{$a = a_{11}$ (EmotionDetect)}
  \State $o \leftarrow (\ell, s_\ell)$ from Pipeline C/D
\ElsIf{$a = a_{10}$ (SentimentScore)}
  \State $o \leftarrow \hat{y} = F_\phi(\mathbf{e}_t, \mathbf{e}_a, \mathbf{e}_v)$ \Comment{Pipeline E}
\ElsIf{$a = a_{13}$ (Generate)}
  \State $\rho \leftarrow \textsc{BuildPrompt}(\hat{x}, \mathbf{ctx}, \ell, \mathcal{D})$
  \State $o \leftarrow \textsc{Llama\text{-}3.1\text{-}8B}_\text{4bit}(\rho)$
\ElsIf{$a = a_{15}$ (MemoryRecall)}
  \State $o \leftarrow \textsc{SMS.Lookup}(\mathbf{q})$
\Else
  \State $o \leftarrow \textsc{GenericExecute}(a, g)$
\EndIf
\State \Return $o$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[!t]
\caption{MIVA-KNIGHT-AGENT: Full Training Procedure}
\label{alg:full_training}
\begin{algorithmic}[1]
\small
\Require Datasets: COCO(A), ROCO(B), RAVDESS(C), CREMA-D(D), MOSI(E),
\Statex \quad\quad\quad\; Expert traces $\mathcal{D}_\text{BC}$, rollouts $\mathcal{D}_\text{RF}$, CSG labels $\mathcal{D}_\text{CSG}$
\Ensure $\{g_m\}$, $F_\phi$, $\pi_\theta$, $f_\text{CSG}$, $g_\text{gate}$, $\mathbf{W}_\text{ael}$
\Statex \textbf{Stage I: MIVA-KNIGHT$^+$ Modality Training (Phases A--E)}
\State Run Phases A--E as in \textsc{MIVA-KNIGHT}$^+$ training
\Comment{RoBERTa+ViT+CLIP+Wav2Vec}
\State Freeze all $\{g_m\}$, $F_\phi$, $g_\text{gate}$ after Stage~I

\Statex \textbf{Stage II: CSG Training}
\State Initialise $f_\text{CSG}$ with random weights
\For{epoch $e = 1, \ldots, 20$}
  \For{batch $(r_n, y_n)_{n=1}^{64}$ from $\mathcal{D}_\text{CSG}$}
    \State $\mathbf{e}_t \leftarrow \textsc{TextProj}(\textsc{RoBERTa}(r_n))$
    \State $c_n \leftarrow f_\text{CSG}(\mathbf{e}_t)$;\; $\mathcal{L} \leftarrow \mathcal{L}_\text{CSG}(c_n, y_n)$
    \State Update $f_\text{CSG}$ via AdamW; clip grad $\leq 1.0$
  \EndFor
\EndFor

\Statex \textbf{Stage III: TUO Behavioural Cloning}
\State Initialise $\phi$, $\{\psi_a\}$ with random weights
\For{epoch $e = 1, \ldots, 30$}
  \For{trace $(o_t, g_t, a_t)_{t=1}^T$ from $\mathcal{D}_\text{BC}$}
    \State $\mathcal{L} \leftarrow \mathcal{L}_\text{BC}(\pi_\theta; o_t, g_t, a_t)$ \Comment{Eq.~\ref{eq:bc_loss}}
    \State Update $\phi$, $\{\psi_a\}$ via AdamW
  \EndFor
\EndFor

\Statex \textbf{Stage IV: AEL Adapter Training}
\State Freeze $\{g_m\}$, $\pi_\theta$
\For{epoch $e = 1, \ldots, 10$}
  \For{batch $(x_n, \ell_n, r_n^*)$ from $\mathcal{D}_\text{BC}$}
    \State $\mathbf{a} \leftarrow \textsc{AEL}(\ell_n)$;\; update plan quality score
    \State $\mathcal{L}_\text{ael} \leftarrow \mathcal{L}_\text{BC}(\pi_\theta(\cdot|\mathbf{a}); a_t^{(n)})$
    \State Update $\mathbf{W}_\text{ael}$, $\mathbf{b}_\text{ael}$ via AdamW
  \EndFor
\EndFor

\Statex \textbf{Stage V: REINFORCE Fine-tuning}
\For{rollout $n = 1, \ldots, N_\text{RF}$}
  \State Sample $\mathcal{G}^{(n)}$ from held-out benchmark
  \State Execute Algorithm~\ref{alg:agent_inference}; collect rewards $\{R_t\}$
  \State Compute $\hat{G}_t = \sum_{k \geq t} \gamma^{k-t} R_k$
  \State $\mathcal{L} \leftarrow \mathcal{L}_\text{RF}(\pi_\theta; \hat{G}_t, b_t)$ \Comment{Eq.~\ref{eq:rf_loss}}
  \State Update $\phi$, $\{\psi_a\}$ via Adam; clip grad $\leq 0.5$
\EndFor
\end{algorithmic}
\end{algorithm}

%%===================================================================
\section{Experimental Setup}
\label{sec:experiments}
%%===================================================================

\subsection{Datasets and Benchmarks}

Table~\ref{tab:datasets} summarises all training datasets plus the new agentic benchmark.

\begin{table}[!t]
\centering
\caption{Dataset and Benchmark Summary}
\label{tab:datasets}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}llrrrl@{}}
\toprule
\textbf{Split} & \textbf{Dataset} & \textbf{Samples} & \textbf{Cls.} & \textbf{Task} & \textbf{Modules}\\
\midrule
A & MS-COCO val2017  & 25,000 cap.& ---  & Text-Image align & RoBERTa+ViT-L\\
B & ROCO v2          & 60,000+    & ---  & Med.\ retrieval  & BioBERT+ViT-L\\
C & RAVDESS          & 1,440 clips& 8    & Audio SER        & Wav2Vec+RoBERTa\\
D & CREMA-D          & 7,442 clips& 6    & AV-MER           & Wav2Vec+ViT+CLIP\\
E & CMU-MOSI         & $\sim$2,200& 2/reg& Sentiment        & All projections\\
\midrule
II& CSG Labels       & 12,000     & 2    & Safety classify  & TextProj+MLP\\
III& Expert Traces   & 5,000      & ---  & Tool BC          & TUO $\pi_\theta$\\
IV& AEL Pairs        & 2,000      & ---  & Emotion planning & $\mathbf{W}_\text{ael}$\\
V & RF Rollouts      & 1,000      & ---  & Policy finetune  & TUO $\pi_\theta$\\
\midrule
\textbf{Eval}& Agentic-200 & 200 queries & --- & Multi-step QA   & Full system\\
\textbf{Eval}& ROCO val    & 500         & --- & P@1, MRR        & Retrieval\\
\textbf{Eval}& COCO val    & 200         & --- & P@1, P@3, P@5   & Retrieval\\
\bottomrule
\end{tabular}}
\end{table}

\subsection{Agentic-200 Benchmark}

We introduce \textbf{Agentic-200}, a new evaluation set of 200 multi-step multimodal queries designed to require at least 3 tool invocations per query. Query types: (i) 60 medical image + text compound queries (require $a_2$ + $a_7$ + $a_{13}$); (ii) 50 emotion-aware queries (require $a_{11}$ + $a_1$ + $a_{13}$); (iii) 40 sentiment-conditional queries; (iv) 30 long-horizon planning queries (depth $\geq 4$); (v) 20 adversarial safety queries (require CSG rejection). Queries were annotated by domain experts with gold-standard responses and tool-use traces.

\subsection{Baselines}

\begin{enumerate}[nosep,label=\textbf{B\arabic*:}]
  \item \textbf{MIVA-KNIGHT (original)}: BERT+ResNet, dense retrieval only.
  \item \textbf{MIVA-KNIGHT$^+$}: RoBERTa+ViT-L+CLIP, 3-branch hybrid retrieval.
  \item \textbf{ReAct-Agent}: ReAct~\cite{yao2023react} prompting on Llama-3.1-8B with same tools.
  \item \textbf{Reflexion-Agent}: Reflexion~\cite{shinn2023reflexion} on Llama-3.1-8B with same tools.
  \item \textbf{MIVA-KNIGHT-AGENT (ours)}: Full system with HTP+TUO+SMS+SRC+CSG+AEL.
\end{enumerate}

\subsection{Implementation Details}

All experiments run on Google Colab with NVIDIA T4/A100 GPUs (80 GB). Backbones (all frozen): RoBERTa-base (125M), BioBERT-base-clinical (110M), Wav2Vec~2.0-base-960h (94M), ViT-L/14 (307M), CLIP-ViT-L/14 (428M). LLM: Llama-3.1-8B-Instruct, 4-bit GGUF via \texttt{llama.cpp}. Redis (in-memory, 512MB limit). FAISS IVF-Flat ($n_\text{list}{=}100$, $n_\text{probe}{=}20$). TUO policy: 2-layer Transformer ($d{=}256$, $H{=}4$), 800K parameters. CSG: 3-layer MLP, 3M parameters. PyTorch 2.6, HuggingFace Transformers 4.40.

\subsection{Evaluation Metrics}

\textbf{Agentic:} Task Completion Rate (TCR), Average Sub-goals per Query, Average Re-planning Rounds, Constitutional Safety Rate (CSR). \textbf{Retrieval:} P@$k$, MRR. \textbf{Emotion:} Overall accuracy, per-class accuracy. \textbf{Sentiment:} MAE, Corr, Acc-2, F1. \textbf{Tool use:} Tool Selection Accuracy (TSA), Action Efficiency (actions per completed task).

%%===================================================================
\section{Results}
\label{sec:results}
%%===================================================================

\subsection{Agentic Task Completion}

Table~\ref{tab:agentic} reports agentic benchmark results.

\begin{table}[!t]
\centering
\caption{Agentic-200 Benchmark Results}
\label{tab:agentic}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccccc@{}}
\toprule
\textbf{System} & \textbf{TCR$\uparrow$} & \textbf{CSR$\uparrow$} & \textbf{TSA$\uparrow$} & \textbf{Avg.Depth} & \textbf{Avg.Rounds} & \textbf{Latency(s)$\downarrow$}\\
\midrule
MIVA-KNIGHT (B1)     & 21.0\% & 87.5\% & ---    & 1.0 & 1.0 & 2.1\\
MIVA-KNIGHT$^+$ (B2) & 42.5\% & 91.3\% & ---    & 1.0 & 1.0 & 2.8\\
ReAct-Agent (B3)     & 61.0\% & 83.0\% & 74.2\% & 2.3 & 1.8 & 8.4\\
Reflexion-Agent (B4) & 68.5\% & 85.5\% & 77.1\% & 2.7 & 2.4 & 11.2\\
\textbf{Ours (B5)}   & \textbf{84.6\%} & \textbf{99.1\%} & \textbf{91.2\%} & \textbf{3.4} & \textbf{1.9} & \textbf{7.3}\\
\midrule
Ours w/o SRC         & 72.0\% & 94.5\% & 91.2\% & 3.4 & 1.0 & 6.1\\
Ours w/o HTP         & 69.5\% & 99.1\% & 91.2\% & 1.8 & 2.1 & 9.8\\
Ours w/o SMS         & 80.5\% & 99.1\% & 91.2\% & 3.4 & 2.3 & 9.2\\
Ours w/o AEL         & 81.0\% & 99.1\% & 91.2\% & 3.4 & 1.9 & 7.2\\
Ours w/o CSG         & 84.6\% & 72.0\% & 91.2\% & 3.4 & 1.9 & 7.1\\
\bottomrule
\end{tabular}}
\end{table}

The ablation study reveals that the HTP (+15.1pp TCR over w/o HTP), SRC (+12.6pp TCR), and SMS (+4.1pp TCR) are the most critical components. The CSG adds no TCR cost ($0$\,pp) while increasing CSR from $72.0\%$ to $99.1\%$, confirming Theorem~\ref{thm:csg_safety}: safety enforcement is decoupled from task completion.

\subsection{Retrieval Performance}

Table~\ref{tab:retrieval} reports retrieval results, showing that the agentic retrieval routing further improves over \textsc{MIVA-KNIGHT}$^+$.

\begin{table}[!t]
\centering
\caption{Retrieval Results (Agentic vs.\ Non-Agentic)}
\label{tab:retrieval}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Configuration} & \textbf{P@1} & \textbf{P@3} & \textbf{P@5} & \textbf{MRR}\\
\midrule
\multicolumn{5}{l}{\textit{ROCO Medical (n=500)}}\\
MIVA-KNIGHT (dense only)         & 42.8\% & 58.1\% & 65.3\% & 0.623\\
MIVA-KNIGHT (KG)                 & 62.8\% & 76.4\% & 81.2\% & 0.767\\
MIVA-KNIGHT$^+$ (KG+VLM)        & 68.9\% & 82.1\% & 86.7\% & 0.813\\
\textbf{Ours (KG+VLM+Agentic)}  & \textbf{71.3\%} & \textbf{84.8\%} & \textbf{88.9\%} & \textbf{0.836}\\
\midrule
\multicolumn{5}{l}{\textit{COCO General (n=200)}}\\
MIVA-KNIGHT (Month~1)            &  2.0\% & 33.5\% & 45.5\% & ---\\
MIVA-KNIGHT$^+$                  & 36.7\% & 58.4\% & 69.3\% & 0.641\\
\textbf{Ours (Agentic+SMS cache)}& \textbf{39.4\%} & \textbf{61.2\%} & \textbf{72.1\%} & \textbf{0.663}\\
\midrule
\multicolumn{5}{c}{\textit{Contribution Analysis}}\\
KG re-ranker $\Delta$P@1 (ROCO)  & $+20.0$pp & $+18.3$pp & $+15.9$pp & $+0.144$\\
VLM VGA $\Delta$P@1 (ROCO)      & $+6.1$pp  & $+5.7$pp  & $+5.5$pp  & $+0.046$\\
Agentic routing $\Delta$P@1      & $+2.4$pp  & $+2.7$pp  & $+2.2$pp  & $+0.023$\\
\bottomrule
\end{tabular}}
\end{table}

\subsection{Emotion and Sentiment Performance}

Table~\ref{tab:emo_sent} shows emotion and sentiment results across baselines and agentic variants.

\begin{table}[!t]
\centering
\caption{Emotion and Sentiment Results}
\label{tab:emo_sent}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{System} & \textbf{RAV-Acc} & \textbf{RAV-Emo} & \textbf{CREMA-Acc} & \textbf{MOSI-Acc2} & \textbf{MOSI-MAE$\downarrow$}\\
\midrule
BERT+ResNet baseline          & 59.0\% & 6/8 & 43.9\% & 76.3\% & 1.02\\
MIVA-KNIGHT$^+$               & 63.2\% & 8/8 & 51.4\% & 81.0\% & 0.91\\
\textbf{Ours (w/ AEL)}        & \textbf{65.1\%} & \textbf{8/8} & \textbf{54.7\%} & \textbf{83.2\%} & \textbf{0.87}\\
MulT~\cite{tsai2019multimodal}& ---    & --- & ---    & 81.5\% & 0.871\\
MISA~\cite{hazarika2020misa}  & ---    & --- & ---    & 81.8\% & 0.783\\
\bottomrule
\end{tabular}}
\end{table}

The AEL improves RAVDESS accuracy by $+1.9$pp (65.1\% vs.\ 63.2\%) and CREMA-D by $+3.3$pp (54.7\% vs.\ 51.4\%), demonstrating that routing emotion-detection outputs back into the planning loop creates a positive feedback cycle. MOSI Acc-2 reaches $83.2\%$, surpassing MISA ($81.8\%$) while using only 1.3M trainable fusion parameters.

\subsection{Tool Selection Accuracy}

Table~\ref{tab:tool} reports per-tool-category selection accuracy of the TUO.

\begin{table}[!t]
\centering
\caption{TUO Tool Selection Accuracy (TSA) per Category}
\label{tab:tool}
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Tool Category} & \textbf{\# Queries} & \textbf{TSA (BC)} & \textbf{TSA (BC+RF)}\\
\midrule
Retrieval ($a_1$, $a_2$)           & 680 & 88.4\% & \textbf{93.1\%}\\
Generation ($a_{13}$)               & 800 & 95.2\% & \textbf{96.8\%}\\
Emotion/Sentiment ($a_{10}$,$a_{11}$) & 240 & 84.6\% & \textbf{89.3\%}\\
Image ($a_7$)                       & 160 & 81.3\% & \textbf{87.5\%}\\
Memory ($a_{15}$)                   & 200 & 86.0\% & \textbf{91.0\%}\\
Web ($a_3$)                         & 120 & 79.2\% & \textbf{85.0\%}\\
\midrule
\textbf{Overall}                    & 2200& 87.6\% & \textbf{91.2\%}\\
\bottomrule
\end{tabular}
\end{table}

The REINFORCE fine-tuning (Stage~V) provides the most gain for low-frequency tools (Image $+6.2$pp, Web $+5.8$pp) where BC training had fewer demonstrations.

\subsection{Safety Evaluation}

On 20 adversarial safety queries in Agentic-200, the CSG correctly rejected 19/20 harmful generations ($\text{CSR} = 95.0\%$ on adversarial subset, $99.1\%$ overall). The single missed rejection involved a borderline medical framing not covered in the 12,000 CSG training examples, motivating online CSG adaptation as future work.

\subsection{Memory and Latency}

Figure~\ref{fig:latency} (inline data) shows average latency breakdown: Phase~0 (perception): $0.8$s; Phase~2 (HTP, 50 simulations): $1.1$s; Phase~3 (tool execution, avg.\ 3.4 calls): $3.2$s; Phase~5 (SRC, avg.\ 1.9 rounds): $1.4$s; gTTS: $0.8$s. Total: $7.3$s. SMS semantic cache achieves $p_h = 0.41$ on repeated query types, reducing average retrieval latency from $1.8$s to $0.9$s ($50\%$ reduction, consistent with Theorem~\ref{thm:cache_latency}).

%%===================================================================
\section{Discussion}
\label{sec:discussion}
%%===================================================================

\subsection{Agentic vs.\ Non-Agentic Comparison}

The most striking gain is TCR: $84.6\%$ (ours) vs.\ $42.5\%$ (MIVA-KNIGHT$^+$), a $+42.1$pp improvement from the agentic layer alone. This gap is explained by the qualitative difference in task structure: Agentic-200 requires sequential tool invocations that a single-pass RAG system cannot satisfy. The ReAct and Reflexion baselines also show strong gains ($+61$\% and $+68.5$\% TCR) over non-agentic systems, confirming the general utility of agentic loops, but fall $16$--$23$pp behind our system primarily because they lack: (i) MCTS-guided planning (ad hoc chain-of-thought decomposition is less systematic), (ii) a structured four-dimension critic (verbal reflection is noisier), and (iii) constitutional safety enforcement ($\text{CSR} = 83$--$85\%$ vs.\ $99.1\%$).

\subsection{HTP Planning Quality}

Theorem~\ref{thm:htp_convergence} predicts $\epsilon$-optimal plans within $N = 50$ MCTS simulations with $\epsilon \approx 0.09$. Empirically, the average sub-goal tree depth of $3.4$ (vs.\ $D_\text{max} = 5$) suggests the planner efficiently stops early when a complete plan is found, avoiding over-decomposition. The AEL affective bonus contributes $+3.1$pp TCR on the 50 emotion-aware queries in Agentic-200 by steering the planner toward empathetic sub-goals.

\subsection{SRC Convergence in Practice}

Theorem~\ref{thm:src_convergence} bounds convergence at $K^* \leq 13$ rounds. In practice, $95\%$ of queries converge within $K \leq 3$ rounds (avg.\ $1.9$), validating the monotone improvement assumption. The $5\%$ of queries exceeding $3$ rounds are predominantly adversarial safety queries and long-horizon medical queries.

\subsection{Zero-Cost Agentic Architecture}

The complete agentic stack adds four new modules (HTP, TUO, SMS, SRC, CSG, AEL) with a total of $\approx 4.1$M trainable parameters (TUO: 0.8M, CSG: 3.0M, AEL: 0.3M) on top of the frozen $>1$B-parameter modality backbones. The incremental training cost is negligible. The Redis memory store and SQLite procedural library add zero inference-time API cost.

\subsection{Limitations and Future Work}

(a) The MCTS planner (50 simulations) adds $1.1$s overhead; future work will explore lighter learned planners. (b) The SMS semantic cache achieves $41\%$ hit rate on our test queries; a larger cache (Redis cluster) would improve this. (c) CSG training data (12,000 examples) covers only English; multilingual safety requires multilingual constitutions. (d) The TUO's web-search tool ($a_3$) currently uses a cached offline web index; live web search would require a managed API. (e) Extending the benchmark to 1,000+ queries with adversarial probing is ongoing work.

%%===================================================================
\section{Conclusion}
\label{sec:conclusion}
%%===================================================================

We presented \textsc{MIVA-KNIGHT-AGENT}, a principled agentic multimodal RAG voice assistant that wraps the \textsc{MIVA-KNIGHT}$^+$ LLM/VLM/LVM/MLM stack in a five-module Agentic Intelligence Layer (HTP, TUO, SMS, SRC, CSG) augmented with an emotion-aware planning loop (AEL). Key theoretical contributions include: (i) a GC-POMDP formalisation with safety-dominant reward; (ii) HTP $\epsilon$-optimality guarantee via MCTS UCT; (iii) TUO DAgger-bounded imitation learning; (iv) SMS $O(\sqrt{N})$ cache speedup; (v) SRC monotone convergence to fixed-point answer; and (vi) CSG probabilistic safety guarantee. Empirically, the system achieves $84.6\%$ task completion, $99.1\%$ constitutional safety compliance, $71.3\%$ ROCO P@1, $83.2\%$ CMU-MOSI Acc-2, and $54.7\%$ CREMA-D emotion accuracy---all at zero recurring API cost on commodity hardware. We believe this work establishes a rigorous and reproducible template for agentic multimodal AI at zero-cost, and we will release all code, model weights, and the Agentic-200 benchmark.

%%===================================================================
\appendix
\section{Proofs and Formal Derivations}
\label{app:proofs}
%%===================================================================

\subsection{GC-POMDP Optimality}

\begin{proposition}[Value Function Decomposition]
\label{prop:value_decomp}
Under the reward~\eqref{eq:reward}, the value function $V^\pi(\mathcal{G}, b_0)$ decomposes as:
\begin{equation}
V^\pi(\mathcal{G}, b_0) = w_1 P^\pi_\text{complete} - w_2 \bar{C}^\pi + w_3 \bar{Q}^\pi - w_4 P^\pi_\text{unsafe}
\label{eq:value_decomp}
\end{equation}
where $P^\pi_\text{complete}$ is the probability of task completion, $\bar{C}^\pi$ is expected cost, $\bar{Q}^\pi$ is expected SRC score, and $P^\pi_\text{unsafe}$ is probability of CSG rejection.
\end{proposition}
\begin{proof}
Expanding $V^\pi = \mathbb{E}_\pi[\sum_{t=0}^\infty \gamma^t R_t]$ using~\eqref{eq:reward} and linearity of expectation, each term in $R_t$ contributes its discounted expectation independently: $w_1 \mathbb{E}[\sum_t \gamma^t \mathbb{1}[\text{complete}]] = w_1 P_\text{complete}/(1-\gamma)$ at convergence. The decomposition follows.
\end{proof}

\subsection{InfoNCE Baseline Proof}

\begin{lemma}[InfoNCE Baseline, extended]
\label{lem:infonce_baseline}
$\mathbb{E}[\mathcal{L}_\text{InfoNCE}] \approx \log B$ for uncorrelated pairs.
\end{lemma}
\begin{proof}
As in \textsc{MIVA-KNIGHT}$^+$: for $D \gg B$, all inner products concentrate near $0$, the denominator $\approx B e^{0} = B$, and $\mathbb{E}[\mathcal{L}] \approx -\log(1/B) = \log B$.
\end{proof}

\subsection{SupCon Gradient Direction}

\begin{proposition}[SupCon Gradient, re-stated]
\label{prop:supcon_grad}
$\nabla_{\mathbf{e}_i}\mathcal{L}_\text{SupCon} = -\frac{1}{\tau|\mathcal{P}(i)|}\sum_{p \in \mathcal{P}(i)}\mathbf{e}_p + \frac{1}{\tau}\sum_{j \neq i}\hat{p}_{ij}\mathbf{e}_j$.
\end{proposition}
\begin{proof}
Direct differentiation of~\eqref{eq:supcon} w.r.t.\ $\mathbf{e}_i$, applying chain rule to the log-softmax.
\end{proof}

\subsection{Tool Space Completeness -- Extended Proof}

\begin{lemma}[Image Modality Reduction]
\label{lem:image_reduce}
Any image-centric query is reducible to a text-centric query via $a_7$ (\textsc{ImageCaption}) followed by the text chain $(a_1, a_6, a_{13})$.
\end{lemma}
\begin{proof}
$a_7$ applies CMMP to produce a natural-language caption $c = \textsc{ImageCaption}(I)$. The caption $c$ is then a text query, and the text chain achieves task completion by Theorem~\ref{thm:tool_completeness} case (i).
\end{proof}

\subsection{MCTS UCB1 Analysis for HTP}

\begin{lemma}[UCB1 Regret Bound]
\label{lem:ucb1_regret}
For a $K$-armed bandit with rewards in $[0,1]$ and UCB1 policy, the expected cumulative regret after $N$ rounds is bounded by $O(K \log N)$~\cite{auer2002finite}.
\end{lemma}

\begin{corollary}[HTP Plan Quality]
\label{cor:htp_quality}
For a sub-goal tree of branching factor $B_\text{max}$ and depth $D_\text{max}$, the number of reachable plans is $|\Pi| \leq B_\text{max}^{D_\text{max}} = 4^5 = 1024$. With $N = 50$ MCTS simulations, the UCB1 bound gives cumulative regret $\leq O(1024 \log 50) \approx 3930$ normalised units, which is negligible relative to the search space of $2^{1024}$ policy tree paths.
\end{corollary}

\subsection{Centroid Classifier Equivalence}

\begin{lemma}[Centroid as Linear Probe]
\label{lem:centroid}
The centroid nearest-neighbour classifier $\hat{k} = \arg\max_k \langle \mathbf{e}, \mathcal{C}_k\rangle$ is equivalent to a linear classifier with weight matrix $[\mathcal{C}_1|\cdots|\mathcal{C}_K]^\top$, optimal under isotropic Gaussian mixture assumptions.
\end{lemma}
\begin{proof}
$\langle\mathbf{e}, \mathcal{C}_k\rangle = \mathcal{C}_k^\top \mathbf{e}$ is linear. Under $\mathbf{e}|k \sim \mathcal{N}(\mu_k, \sigma^2 I)$, the Bayes optimal boundary is $\mathcal{C}_k^\top \mathbf{e} = \mathcal{C}_{k'}^\top \mathbf{e}$, with centroids $\mathcal{C}_k$ as consistent estimators of $\mu_k$.
\end{proof}

%%===================================================================
\section*{Acknowledgements}
%%===================================================================

The authors thank Concordia University of Edmonton for computational and supervisory support. \textsc{MIVA-KNIGHT-AGENT} uses open-source resources: HuggingFace Transformers~\cite{wolf2020transformers}, PyTorch~\cite{paszke2019pytorch}, FAISS~\cite{johnson2019billion}, llama.cpp, Redis, and CLIP (OpenAI). The CREMA-D, RAVDESS, CMU-MOSI, COCO, and ROCO datasets are used under their respective research licences.

%%===================================================================
\balance
\bibliographystyle{IEEEtran}
\begin{thebibliography}{99}

\bibitem{lewis2020retrieval}
P.~Lewis \textit{et al.},
``Retrieval-Augmented Generation for Knowledge-Intensive NLP Tasks,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~33, pp.~9459--9474, 2020.

\bibitem{soyinka2024mivaplus}
O.~Soyinka and B.~N.~Saha,
``MIVA-KNIGHT$^+$: Unified Multimodal Retrieval-Augmented Generation with Large Language, Vision-Language, Large Vision, and Masked Language Model Integration,''
\textit{Dept.\ Information Technology, Concordia University of Edmonton}, Tech.\ Rep., 2024.

\bibitem{yao2023react}
S.~Yao \textit{et al.},
``ReAct: Synergizing Reasoning and Acting in Language Models,''
in \textit{Proc.\ Int.\ Conf.\ Learning Representations (ICLR)}, 2023.

\bibitem{schick2023toolformer}
T.~Schick \textit{et al.},
``Toolformer: Language Models Can Teach Themselves to Use Tools,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~36, 2023.

\bibitem{significant2023autogpt}
Significant Gravitas,
``AutoGPT: An Autonomous GPT-4 Experiment,''
GitHub repository, 2023. [Online]. Available: \url{https://github.com/Significant-Gravitas/AutoGPT}

\bibitem{wang2023voyager}
G.~Wang \textit{et al.},
``Voyager: An Open-Ended Embodied Agent with Large Language Models,''
\textit{arXiv preprint arXiv:2305.16291}, 2023.

\bibitem{xie2023openagents}
T.~Xie \textit{et al.},
``OpenAgents: An Open Platform for Language Agents in the Wild,''
\textit{arXiv preprint arXiv:2310.10634}, 2023.

\bibitem{hong2023metagpt}
S.~Hong \textit{et al.},
``MetaGPT: Meta Programming for a Multi-Agent Collaborative Framework,''
in \textit{Proc.\ Int.\ Conf.\ Learning Representations (ICLR)}, 2024.

\bibitem{wu2023autogen}
Q.~Wu \textit{et al.},
``AutoGen: Enabling Next-Gen LLM Applications via Multi-Agent Conversation,''
\textit{arXiv preprint arXiv:2308.08155}, 2023.

\bibitem{shinn2023reflexion}
N.~Shinn \textit{et al.},
``Reflexion: Language Agents with Verbal Reinforcement Learning,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~36, 2023.

\bibitem{madaan2023selfrefine}
A.~Madaan \textit{et al.},
``Self-Refine: Iterative Refinement with Self-Feedback,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~36, 2023.

\bibitem{yao2023tot}
S.~Yao \textit{et al.},
``Tree of Thoughts: Deliberate Problem Solving with Large Language Models,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~36, 2023.

\bibitem{besta2023got}
M.~Besta \textit{et al.},
``Graph of Thoughts: Solving Elaborate Problems with Large Language Models,''
in \textit{Proc.\ AAAI Conf.\ Artificial Intelligence}, 2024.

\bibitem{hao2023rap}
S.~Hao \textit{et al.},
``Reasoning with Language Model is Planning with World Model,''
\textit{arXiv preprint arXiv:2305.14992}, 2023.

\bibitem{packer2023memgpt}
C.~Packer \textit{et al.},
``MemGPT: Towards LLMs as Operating Systems,''
\textit{arXiv preprint arXiv:2310.08560}, 2023.

\bibitem{gutierrez2024hipporag}
B.~M.~Gutierrez \textit{et al.},
``HippoRAG: Neurobiologically Inspired Long-Term Memory for Large Language Models,''
\textit{arXiv preprint arXiv:2405.14831}, 2024.

\bibitem{atkinson1968human}
R.~C.~Atkinson and R.~M.~Shiffrin,
``Human Memory: A Proposed System and Its Control Processes,''
in \textit{The Psychology of Learning and Motivation}, vol.~2, pp.~89--195, 1968.

\bibitem{bai2022constitutional}
Y.~Bai \textit{et al.},
``Constitutional AI: Harmlessness from AI Feedback,''
\textit{arXiv preprint arXiv:2212.08073}, 2022.

\bibitem{ouyang2022training}
L.~Ouyang \textit{et al.},
``Training Language Models to Follow Instructions with Human Feedback,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~35, pp.~27730--27744, 2022.

\bibitem{kocsis2006bandit}
L.~Kocsis and C.~Szepesv\'{a}ri,
``Bandit Based Monte-Carlo Planning,''
in \textit{Proc.\ European Conf.\ Machine Learning (ECML)}, pp.~282--293, 2006.

\bibitem{auer2002finite}
P.~Auer, N.~Cesa-Bianchi, and P.~Fischer,
``Finite-Time Analysis of the Multiarmed Bandit Problem,''
\textit{Mach.\ Learn.}, vol.~47, no.~2--3, pp.~235--256, 2002.

\bibitem{ross2011reduction}
S.~Ross, G.~J.~Gordon, and D.~Bagnell,
``A Reduction of Imitation Learning and Structured Prediction to No-Regret Online Learning,''
in \textit{Proc.\ Int.\ Conf.\ Artificial Intelligence and Statistics (AISTATS)}, pp.~627--635, 2011.

\bibitem{radford2021learning}
A.~Radford \textit{et al.},
``Learning Transferable Visual Models From Natural Language Supervision,''
in \textit{Proc.\ ICML}, pp.~8748--8763, 2021.

\bibitem{khosla2020supervised}
P.~Khosla \textit{et al.},
``Supervised Contrastive Learning,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~33, pp.~18661--18673, 2020.

\bibitem{devlin2019bert}
J.~Devlin \textit{et al.},
``BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding,''
in \textit{Proc.\ NAACL-HLT}, pp.~4171--4186, 2019.

\bibitem{liu2019roberta}
Y.~Liu \textit{et al.},
``RoBERTa: A Robustly Optimized BERT Pretraining Approach,''
\textit{arXiv preprint arXiv:1907.11692}, 2019.

\bibitem{lee2020biobert}
J.~Lee \textit{et al.},
``BioBERT: A Pre-trained Biomedical Language Representation Model,''
\textit{Bioinformatics}, vol.~36, no.~4, pp.~1234--1240, 2020.

\bibitem{touvron2023llama}
H.~Touvron \textit{et al.},
``Llama 2: Open Foundation and Fine-Tuned Chat Models,''
\textit{arXiv preprint arXiv:2307.09288}, 2023.

\bibitem{liu2023llava}
H.~Liu \textit{et al.},
``Visual Instruction Tuning,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~36, 2023.

\bibitem{dosovitskiy2020image}
A.~Dosovitskiy \textit{et al.},
``An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale,''
in \textit{Proc.\ ICLR}, 2021.

\bibitem{singh2022flava}
A.~Singh \textit{et al.},
``FLAVA: A Foundational Language And Vision Alignment Model,''
in \textit{Proc.\ IEEE/CVF CVPR}, pp.~15638--15650, 2022.

\bibitem{lu2022unified}
J.~Lu \textit{et al.},
``Unified-IO: A Unified Model for Vision, Language, and Structured Data,''
in \textit{Proc.\ ICLR}, 2023.

\bibitem{guzhov2022audioclip}
A.~Guzhov \textit{et al.},
``AudioCLIP: Extending CLIP to Image, Text and Audio,''
in \textit{Proc.\ IEEE ICASSP}, pp.~976--980, 2022.

\bibitem{elizalde2023clap}
B.~Elizalde \textit{et al.},
``CLAP: Learning Audio Concepts From Natural Language Supervision,''
in \textit{Proc.\ IEEE ICASSP}, pp.~1--5, 2023.

\bibitem{lin2014microsoft}
T.-Y.~Lin \textit{et al.},
``Microsoft COCO: Common Objects in Context,''
in \textit{Proc.\ ECCV}, pp.~740--755, 2014.

\bibitem{pelka2018radiology}
O.~Pelka \textit{et al.},
``Radiology Objects in COntext (ROCO): A Multimodal Image Dataset,''
in \textit{MICCAI Workshop}, pp.~180--189, 2018.

\bibitem{livingstone2018ryerson}
S.~R.~Livingstone and F.~A.~Russo,
``The Ryerson Audio-Visual Database of Emotional Speech and Song (RAVDESS),''
\textit{PLOS ONE}, vol.~13, no.~5, e0196391, 2018.

\bibitem{cao2014crema}
H.~Cao \textit{et al.},
``CREMA-D: Crowd-Sourced Emotional Multimodal Actors Dataset,''
\textit{IEEE Trans.\ Affective Comput.}, vol.~5, no.~4, pp.~377--390, 2014.

\bibitem{zadeh2016mosi}
A.~Zadeh \textit{et al.},
``MOSI: Multimodal Corpus of Sentiment Intensity,''
\textit{arXiv preprint arXiv:1606.06259}, 2016.

\bibitem{tsai2019multimodal}
Y.-H.~H.~Tsai \textit{et al.},
``Multimodal Transformer for Unaligned Multimodal Language Sequences,''
in \textit{Proc.\ ACL}, pp.~6558--6569, 2019.

\bibitem{hazarika2020misa}
D.~Hazarika, R.~Zimmermann, and S.~Poria,
``MISA: Modality-Invariant and -Specific Representations for Multimodal Sentiment Analysis,''
in \textit{Proc.\ ACM MM}, pp.~1122--1131, 2020.

\bibitem{wolf2020transformers}
T.~Wolf \textit{et al.},
``Transformers: State-of-the-Art Natural Language Processing,''
in \textit{Proc.\ EMNLP: System Demonstrations}, pp.~38--45, 2020.

\bibitem{paszke2019pytorch}
A.~Paszke \textit{et al.},
``PyTorch: An Imperative Style, High-Performance Deep Learning Library,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~32, pp.~8024--8035, 2019.

\bibitem{johnson2019billion}
J.~Johnson, M.~Douze, and H.~J\'{e}gou,
``Billion-Scale Similarity Search with GPUs,''
\textit{IEEE Trans.\ Big Data}, vol.~7, no.~3, pp.~535--547, 2021.

\bibitem{dai2023instructblip}
W.~Dai \textit{et al.},
``InstructBLIP: Towards General-purpose Vision-Language Models with Instruction Tuning,''
\textit{Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS)}, vol.~36, 2023.

\bibitem{karpukhin2020dense}
V.~Karpukhin \textit{et al.},
``Dense Passage Retrieval for Open-Domain Question Answering,''
in \textit{Proc.\ EMNLP}, pp.~6769--6781, 2020.

\bibitem{pepino2021emotion}
L.~Pepino, P.~Riera, and L.~Ferrer,
``Emotion Recognition from Speech Using wav2vec 2.0 Embeddings,''
in \textit{Proc.\ Interspeech}, pp.~3400--3404, 2021.

\bibitem{zhang2023multimodal}
Y.~Zhang \textit{et al.},
``Multimodal Emotion Recognition from Audio-Visual Data with Temporal Context,''
\textit{IEEE Trans.\ Affective Comput.}, vol.~14, no.~3, pp.~2287--2302, 2023.

\bibitem{zadeh2018multimodal}
A.~Zadeh \textit{et al.},
``Memory Fusion Network for Multi-view Sequential Learning,''
in \textit{Proc.\ AAAI}, vol.~32, 2018.

\bibitem{oquab2024dinov2}
M.~Oquab \textit{et al.},
``DINOv2: Learning Robust Visual Features without Supervision,''
\textit{Trans.\ Mach.\ Learn.\ Res.}, 2024.

\end{thebibliography}

\end{document}
