% ============================================================
%  MIVA-KNIGHT — Refined Abstract
%  Multi-modal Industrial Voice Assistant with
%  Knowledge-Integrated Hybrid Reasoning (Hybrid RAG)
% ============================================================
\documentclass[12pt,a4paper]{article}

% ── Core packages ──────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames,svgnames]{xcolor}
\usepackage[most]{tcolorbox}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{hyperref}

% ── Hyperref setup ─────────────────────────────────────────
\hypersetup{
  colorlinks=true,
  linkcolor=NavyBlue,
  citecolor=ForestGreen,
  urlcolor=Maroon,
  pdftitle={MIVA-KNIGHT Abstract},
  pdfauthor={MIVA-KNIGHT Project}
}

% ── Colour palette ─────────────────────────────────────────
\definecolor{mivablue}{RGB}{25,84,163}
\definecolor{mivagreen}{RGB}{34,139,34}
\definecolor{mivaorange}{RGB}{210,105,30}
\definecolor{mivapurple}{RGB}{106,13,173}
\definecolor{mivagray}{RGB}{100,100,100}
\definecolor{boxbg}{RGB}{245,248,255}

% ── Section formatting ─────────────────────────────────────
\titleformat{\section}
  {\large\bfseries\color{mivablue}}
  {{\color{mivablue}\thesection}}{0.8em}{}[\titlerule]

% ── Header / Footer ────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\color{mivagray}\textbf{MIVA-KNIGHT} \textemdash\ Abstract}
\fancyhead[R]{\small\color{mivagray}Hybrid RAG $\cdot$ Multi-modal AI}
\fancyfoot[C]{\small\color{mivagray}\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ── tcolorbox style ────────────────────────────────────────
\tcbset{
  abstractsec/.style={
    enhanced, breakable,
    colback=boxbg, colframe=mivablue,
    fonttitle=\bfseries\color{white},
    attach boxed title to top left={yshift=-2mm, xshift=4mm},
    boxed title style={colback=mivablue, rounded corners},
    arc=4pt, outer arc=4pt,
    top=6pt, bottom=6pt, left=8pt, right=8pt
  }
}

% ── Document ───────────────────────────────────────────────
\begin{document}

% ── Title block ────────────────────────────────────────────
\begin{center}
  {\LARGE\bfseries\color{mivablue} MIVA-KNIGHT}\\[0.3cm]
  {\large\color{mivablue!80}
    Multi-modal Industrial Voice Assistant\\[0.15cm]
    with Knowledge-Integrated Hybrid Reasoning}\\[0.5cm]
  {\normalsize\color{mivagray}
    \textit{A Domain-Adaptive, Privacy-Preserving AI Platform\\
    Powered by Hybrid RAG, Structured Knowledge Graphs,\\
    and Explainable Multi-modal Reasoning}}\\[0.6cm]
  \begin{tabular}{cc}
    \toprule
    \textbf{Modalities}      & Text,\ Image,\ Voice,\ Sensor Data \\
    \textbf{Domains}         & Healthcare,\ Energy,\ Manufacturing,\ Finance,\ Education \\
    \textbf{Core Techniques} & Hybrid RAG-LightRAG,\ Knowledge Graph,\ Grad-CAM,\ XAI \\
    \textbf{Deployment}      & On-premise containerised microservices (Docker) \\
    \bottomrule
  \end{tabular}
\end{center}

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

% ════════════════════════════════════════════════════════════
\section*{Background}
% ════════════════════════════════════════════════════════════

Voice assistants have fundamentally changed how humans interact with information, yet
mainstream AI systems continue to prioritise surface-level answer generation over
factual depth and verifiability. Trained predominantly on broad, noisy internet corpora,
today's large language models (LLMs) exhibit strong general reasoning but fail critically
when applied to privacy-sensitive industrial domains that demand rigorously grounded
knowledge---such as medical diagnostics, corporate governance, or industrial process control.

Three structural deficiencies limit the adoption of current architectures in these sectors.
First, \textbf{lack of structured knowledge grounding}: models are unaware of
expert-curated domain ontologies, making them prone to confident hallucination.
Second, \textbf{unimodal perception}: state-of-the-art systems rarely integrate text,
images, audio, and sensor streams in a unified, semantically coherent framework.
Third, \textbf{cloud dependency and privacy risk}: reliance on remote APIs makes
deployment in privacy-critical environments---hospitals, energy infrastructure,
financial institutions---operationally and legally untenable.

Alberta's diversified industrial economy provides both the urgent motivation and the
real-world testbeds required to address these challenges at scale.

% ════════════════════════════════════════════════════════════
\section*{Purpose}
% ════════════════════════════════════════════════════════════

This project develops \textbf{MIVA-KNIGHT}
(\textbf{M}ulti-modal \textbf{I}ndustrial \textbf{V}oice \textbf{A}ssistant with
\textbf{K}nowledge-\textbf{N}etwork \textbf{I}ntegrated \textbf{G}raph \textbf{H}ybrid
\textbf{T}ransformer)---an end-to-end, domain-adaptive AI platform designed to shift the
paradigm from \emph{``knowing it all''} to \emph{``knowing it right''}.

MIVA-KNIGHT grounds every generated response in verifiable, structured domain evidence
by fusing a Hybrid Retrieval-Augmented Generation (RAG-LightRAG) framework with
Neo4j-based domain Knowledge Graphs, multi-modal transformer fusion, and visual
explainability via Grad-CAM. All components operate entirely on-premise, within
containerised microservices, to meet the strictest institutional data-sovereignty
requirements.

% ════════════════════════════════════════════════════════════
\section*{Methods}
% ════════════════════════════════════════════════════════════

\begin{tcolorbox}[abstractsec, title={Architecture Overview}]
MIVA-KNIGHT is the functional composition of five independent sub-systems:
\[
  \text{MIVA-KNIGHT}
  \;=\; S_{\mathrm{KG}} \;\circ\; S_{\mathrm{Retrieval}}
  \;\circ\; S_{\mathrm{Fusion}} \;\circ\; S_{\mathrm{Reasoning}}
  \;\circ\; S_{\mathrm{Generator}}
\]
Each sub-system exposes well-defined input/output contracts and can be swapped
independently. If any input modality is unavailable, its embedding is replaced by
the zero vector $e_m = \mathbf{0}_{d_m}$, ensuring the pipeline degrades
gracefully without failure.
\end{tcolorbox}

\textbf{Input processing.}
The system accepts text, images, voice, and numeric sensor streams simultaneously.
Dedicated encoders---BERT for text, a ViT/ResNet encoder for images, Wav2Vec~2.0
for audio, and a lightweight MLP for sensor time-series---map each modality into a
shared 512-dimensional embedding space via learnable linear projection heads. This
design keeps foundation encoders frozen across domains while allowing rapid
domain adaptation by updating only the projection heads.

\textbf{Hybrid Retrieval (RAG-LightRAG).}
Candidate evidence is retrieved via a two-stage hybrid mechanism. The \emph{dense
retrieval} stage computes cosine similarity between the fused query embedding and a
FAISS vector index of domain documents:
\[
  \mathrm{score}_{\mathrm{dense}}(q, d)
  = \frac{\mathrm{Enc}(q)\cdot\mathrm{Enc}(d)}
         {\|\mathrm{Enc}(q)\|\;\|\mathrm{Enc}(d)\|}
\]
The \emph{graph retrieval} stage (LightRAG) performs multi-hop DFS traversal over
a Neo4j Knowledge Graph $\mathcal{G} = (\mathcal{V}, \mathcal{E}, \mathcal{R})$,
scoring paths by inverse length:
\[
  \mathrm{score}_{\mathrm{graph}}(q, e)
  = \sum_{p \in \mathcal{P}(q,\,e)} \frac{1}{|p|+1}\cdot\mathrm{rel}(p)
\]
The two scores are combined via a convex mixture controlled by $\alpha \in [0,1]$:
\[
  \boxed{
    \mathrm{score}_{\mathrm{hybrid}}(q, c)
    = \alpha\cdot\mathrm{score}_{\mathrm{dense}}(q, c)
    + (1-\alpha)\cdot\mathrm{score}_{\mathrm{graph}}(q, c)
  }
\]
Setting $\alpha = 1$ recovers pure dense retrieval; $\alpha = 0$ recovers pure
graph retrieval.

\textbf{Multi-modal fusion.}
Projected modality embeddings are stacked into a sequence
$\mathbf{X} = [h_t;\, h_i;\, h_v;\, h_s]$ and passed through an 8-head
Multi-Head Attention fusion layer with residual connections and layer normalisation:
\[
  e_{\mathrm{fused}}
  = \mathrm{MeanPool}\!\Bigl(
      \mathrm{LN}\!\bigl[\mathbf{X}' + \mathrm{FFN}(\mathbf{X}')\bigr]
    \Bigr),
  \quad
  \mathbf{X}' = \mathrm{LN}\!\bigl(\mathbf{X} + \mathrm{MHA}(\mathbf{X})\bigr)
\]
Cross-modal alignment is enforced during training via a contrastive loss (InfoNCE)
targeting cosine similarity $\geq 0.8$ between paired modalities.

\textbf{Knowledge Graph grounding.}
A Named Entity Recognition (NER) module extracts entities from the query; the KG
module retrieves multi-hop relational context $\mathbf{K}_{\mathrm{KG}}$ via DFS
up to depth $k = 2$. Domain graphs are constructed in Neo4j with $10{,}000$--$50{,}000$
entities per domain, spanning healthcare, energy, manufacturing, finance, and education.

\textbf{Transformer reasoning and response generation.}
A 4-layer Transformer Encoder processes $e_{\mathrm{fused}}$ with cross-attention
over $\mathbf{K}_{\mathrm{KG}}$:
\[
  r = \mathrm{OutputHead}\!\Bigl(
        x + \mathrm{Attention}(x,\, \mathbf{K}_{\mathrm{KG}},\, \mathbf{K}_{\mathrm{KG}})
      \Bigr),
  \quad
  x = \mathrm{TransformerEncoder}^{(4)}(e_{\mathrm{fused}})
\]
A confidence-tiered mechanism rejects low-confidence queries before generation to
prevent hallucination. The final response $\hat{y} = \mathrm{FFN}(r)$ is accompanied
by a Grad-CAM visual explanation localising the image regions that drove the answer:
\[
  \mathrm{CAM}^c
  = \mathrm{ReLU}\!\left(\sum_k \alpha_k^c\, A^k\right),
  \quad
  \alpha_k^c = \frac{1}{Z}\sum_{i,j}
    \frac{\partial y^c}{\partial A_{ij}^k}
\]

\textbf{Training objective.}
The system is trained end-to-end with a multi-objective loss:
\[
  \mathcal{L}
  = \mathcal{L}_{\mathrm{task}}
  + \lambda_1\,\mathcal{L}_{\mathrm{contr}}
  + \lambda_2\,\mathcal{L}_{\mathrm{KG}}
  + \lambda_3\,\mathcal{L}_{\mathrm{explain}}
\]

\textbf{Deployment.}
All components are containerised via Docker as independent microservices
(API Gateway, Voice Service, Retrieval Service, Knowledge Graph Service, Reasoning Service,
Explanation Service) communicating through message queues (RabbitMQ/Kafka) with Redis
caching---enabling secure, fully offline, on-premise operation.

% ════════════════════════════════════════════════════════════
\section*{Anticipated Results}
% ════════════════════════════════════════════════════════════

Quantitative targets are established across five evaluation dimensions:

\begin{center}
\begin{tabular}{lcc}
\toprule
\textbf{Metric} & \textbf{Formula} & \textbf{Target} \\
\midrule
Accuracy          & $|\hat{y}=y|/N$                       & $\geq 90\%$    \\
Macro-F1          & $\overline{F1_c}$ across classes       & $\geq 0.88$    \\
MRR               & $\frac{1}{|Q|}\sum_i 1/r_i$           & $\geq 0.80$    \\
Precision@1       & Top-1 retrieval accuracy              & $> 60\%$       \\
Precision@3       & Top-3 retrieval accuracy              & $> 75\%$       \\
Interpretability  & $1 - H(A)/\log|A|$                    & $\geq 0.60$    \\
Inference latency & $\mathbb{E}[t]$ over 100 trials       & $< 2\,\mathrm{s}$ \\
Domain hot-swap   & FAISS index replacement time          & $< 5\,\mathrm{s}$ \\
\bottomrule
\end{tabular}
\end{center}

The prototype is expected to demonstrate \textbf{zero-shot domain adaptability}---
seamlessly transitioning between specialisations (e.g.\ medical radiology and corporate
governance) by hot-swapping only the FAISS index and domain Knowledge Graph, while
the shared multi-modal embedding space and transformer backbone remain intact.
Regression testing will confirm that baseline knowledge is retained without
catastrophic forgetting.

Target domain-level impact projections include: $+30\%$ diagnostic accuracy and
$-40\%$ diagnosis time in healthcare; $-15\%$ operational cost in energy;
$-40\%$ equipment downtime in manufacturing; $+25\%$ risk-assessment accuracy
in finance; and $+50\%$ learner engagement with $-60\%$ search overhead in education.

% ════════════════════════════════════════════════════════════
\section*{Conclusions}
% ════════════════════════════════════════════════════════════

MIVA-KNIGHT represents a principled departure from opaque, cloud-dependent AI towards
\textbf{transparent, sovereign, and knowledge-grounded} intelligence. By isolating a
shared multi-modal embedding space from domain-specific FAISS indexes and Knowledge
Graphs, the architecture achieves rapid domain adaptation without retraining foundational
components. By combining hybrid RAG-LightRAG retrieval with structured KG reasoning and
Grad-CAM explainability, it delivers responses that are not only accurate but fully
auditable---every answer is traceable to specific retrieved documents and graph-traversal
paths.

Deployed entirely on-premise within containerised microservices, MIVA-KNIGHT eliminates
the privacy liability inherent in cloud-based inference---a prerequisite for adoption in
healthcare, energy, and financial sectors. The system offers a level of factual
transparency, interpretability, and data sovereignty that no legacy consumer voice
assistant can match.

The open-source, containerised framework will be released to enable continued research
in multi-modal reasoning and explainable AI, positioning Alberta as a leader in
human-centric, sovereign AI innovation.

% ── Bibliography ───────────────────────────────────────────
\vspace{0.5cm}
\begin{thebibliography}{9}
\bibitem{whisper}
  A.~Radford et al., ``Robust Speech Recognition via Large-Scale Weak Supervision,''
  \emph{ICML}, 2023.

\bibitem{bert}
  J.~Devlin et al., ``BERT: Pre-training of Deep Bidirectional Transformers
  for Language Understanding,'' \emph{NAACL}, 2019.

\bibitem{lightrag}
  Edge et al., ``From Local to Global: A Graph RAG Approach to Query-Focused
  Summarization,'' \emph{arXiv:2404.16130}, 2024.

\bibitem{wav2vec}
  A.~Baevski et al., ``wav2vec 2.0: A Framework for Self-Supervised Learning
  of Speech Representations,'' \emph{NeurIPS}, 2020.

\bibitem{gradcam}
  R.~R. Selvaraju et al., ``Grad-CAM: Visual Explanations from Deep Networks
  via Gradient-Based Localization,'' \emph{ICCV}, 2017.

\bibitem{infonce}
  A.~van den Oord et al., ``Representation Learning with Contrastive Predictive
  Coding,'' \emph{arXiv:1807.03748}, 2018.

\bibitem{transformer}
  A.~Vaswani et al., ``Attention Is All You Need,'' \emph{NeurIPS}, 2017.

\bibitem{neo4j}
  I.~Robinson et al., \emph{Graph Databases}, O'Reilly Media, 2015.

\bibitem{faiss}
  J.~Johnson et al., ``Billion-Scale Similarity Search with GPUs,''
  \emph{IEEE Trans.\ Big Data}, 2019.
\end{thebibliography}

\end{document}
